---
title: "Exploratory Analysis"
author: "Maddy Kornhauser, Sabrina Lee, Brian Rawn"
date: "2/4/2021"
output: rmarkdown::github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Class 2/9/2021

## General Updates

* Connected with Andrew last week and are coordinating on some additional data (2019 moves, normalization information).
* Refined our use case.
* Looked into using Yelp data.
* Begun exploratory analysis.

## Use Case

Last week, our use case was focused on the causal effects of nightlife. Not only is this very difficult to measure, there are a number of problems associated with underlying bias in  data we would use to measure negative externalties.

This week, we have split our use case into two parts - one to frame our exploratory analysis, and a second (WIP) to frame the public policy tool we will ultimately buid.

* *Exploratory analysis use case*: How does the location of nightlife establishments impact foot traffic in Philadelphia?

* *Public policy use case*: How far are people willing to travel for nightlife establishments? How much foot traffic can we expect in a given nightlife corridor?

By developing a tool to measure how nightlife establishments impact foot traffic (a proxy for commercial activity), community groups and councils can quickly ascertain the demand for nightlife in certain areas of Philadelphia. This tool will help stakeholders make informed decisions when siting and issuing permits and licenses for future nightlife establishments.

## Yelp data

There is a [Yelp Open Dataset](https://www.yelp.com/dataset) with a subset of businesses, reviews and user data that is available for academic purposes.  Unfortunately, this data is only availabe for a selection of cities that do not include Philadelphia.

Alternatively, we can plug into the Yelp API and pull Philadelphia data meeting various search criteria. Based on online articles (see [here](https://billpetti.github.io/2017-12-23-use-yelp-api-r-rstats/) and [here](https://rpubs.com/fitzpatrickm8/yelpapi)) we wrote a script that can queries the Yelp API. This is also saved on the repo. There are some limitations:

* It appears that you cannot pull more than 50 search results at a time based on a search term.
* It's not clear how search results are ordered.
* How would we match this up against the SafeGraph data?  There is coordinate infromation and names of stores, but we can forsee some issues here.
* Like the 311 and crime data, we have to consider the underying bias of the Yelp data. Is this reliable?

## Exploratory Analysis

This week, we have focused on exploring the 2018 SafeGraph dataset and answering the following questions.

* Where and how far are people travelling from? 
* When are people travelling?
* What is the number of visits/visitors? Is there seasonality in the mobility patterns? Weekly patterns?
* Where does nightlife happen?
* Where are the major nightlife corridors?
* How long do people spend in establishments?  What is the median dwell time?

### Setup

```{r include=FALSE}

library(tidyverse)
library(sf)
library(lubridate)
library(datetime)
library(viridis)
library(gridExtra)
library(dplyr)
library(ggplot2)
library(gganimate)

qBr <- function(df, variable, rnd) {
  if (missing(rnd)) {
    as.character(quantile(round(df[[variable]],0),
                          c(.01,.2,.4,.6,.8), na.rm=T))
  } else if (rnd == FALSE | rnd == F) {
    as.character(formatC(quantile(df[[variable]]), digits = 3),
                 c(.01,.2,.4,.6,.8), na.rm=T)
  }
}
q5 <- function(variable) {as.factor(ntile(variable, 5))}
palette5 <- viridis_pal()(5)
mapTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 14,colour = "black"),
    plot.subtitle=element_text(face="italic"),
    plot.caption=element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),axis.title = element_blank(),
    axis.text = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=2)
  )
}

setwd("~/GitHub/musa_practicum_nighttime")
#setwd("C:/Users/rawnb/Desktop/MUSA/Spring 2021/Practicum")

###########
# LOAD DATA
############
#dat <- read.csv("./moves_2018.csv")
#phila <- st_read("./phila.geojson")

dat <- read.csv("./data/moves_2018.csv")
phila <- st_read("./demo/phila.geojson")

dat2 <- dat %>% 
  dplyr::select(safegraph_place_id, date_range_start, date_range_end, raw_visit_counts,
                       raw_visitor_counts, visits_by_day, poi_cbg, visitor_home_cbgs, 
                       visitor_daytime_cbgs, visitor_work_cbgs, visitor_country_of_origin,
                       distance_from_home, median_dwell, bucketed_dwell_times, related_same_day_brand, 
                       related_same_month_brand, popularity_by_hour, popularity_by_day, device_type) %>%
  # mutate(date_range_start = as.date(dat$date_range_start),
  #        date_range_end = as.date(dat$date_range_end)) %>%
  left_join(., phila, by = "safegraph_place_id") %>% 
  st_as_sf()

phl_cbg <- st_read("http://data.phl.opendata.arcgis.com/datasets/2f982bada233478ea0100528227febce_0.geojson") %>%
  mutate(GEOID10 = as.numeric(GEOID10))
phl_zip <- st_read("http://data.phl.opendata.arcgis.com/datasets/b54ec5210cee41c3a884c9086f7af1be_0.geojson") %>%
  mutate(CODE = as.numeric(CODE))
```

### Where and how far are people travelling from?

The following maps break down the count of different CBGs by point of interest and sectors that contribute to the nighttime economy. Looking first at restaurants, we see that locations that attract the most CBGs are located primarily in Center City and along transit lines. Because of so many restaurants, we have shown a map during summer and winter.

```{r echo=TRUE, message=FALSE, warning=FALSE}
dat_cbg_visitors <- 
  dat2 %>%
  select(safegraph_place_id, date_range_start, top_category, sub_category, poi_cbg, visitor_home_cbgs, geometry) %>%
  mutate(visitor_home_cbgs = str_remove_all(visitor_home_cbgs, pattern = "\\[|\\]")) %>%
  mutate(visitor_home_cbgs = str_remove_all(visitor_home_cbgs, pattern = "\\{|\\}")) %>%
  mutate(visitor_home_cbgs = str_remove_all(visitor_home_cbgs, pattern = '\\"|\\"')) %>%
  mutate(visitor_home_cbgs = str_split(visitor_home_cbgs, pattern = ",")) %>%
  unnest(visitor_home_cbgs) %>%
  separate(.,
           visitor_home_cbgs,
           c("Visitor_CBG", "Visitors"),
           sep = ":") %>%
  mutate(Visitor_CBG = as.numeric(Visitor_CBG),
         poi_cbg = as.numeric(poi_cbg),
         Visitors = as.numeric(Visitors))

#Restaurants
restaurants_cbg_jan <- 
  dat_cbg_visitors %>%
  filter(top_category == "Restaurants and Other Eating Places",
         date_range_start == "2018-01-01T05:00:00Z") %>%
  group_by(safegraph_place_id) %>%
  summarize(Count = n()) %>%
  ggplot() + 
  geom_sf(data = phl_cbg, fill = "grey40", color = "transparent") +
  geom_sf(aes(color = q5(Count)), size = .5) + 
  scale_fill_manual(values = palette5,
                    aesthetics = c("colour", "fill"),
                    name = "CBG Count\nQuintile Breaks") +
  mapTheme() +
  labs(title = "How many different CBGs visit \neach restaurant? (January)")

restaurants_cbg_jul <- 
  dat_cbg_visitors %>%
  filter(top_category == "Restaurants and Other Eating Places",
         date_range_start == "2018-07-01T04:00:00Z") %>%
  group_by(safegraph_place_id) %>%
  summarize(Count = n()) %>%
  ggplot() +
  geom_sf(data = phl_cbg, fill = "grey40", color = "transparent") +
  geom_sf(aes(color = q5(Count)), size = .5) +
  scale_fill_manual(values = palette5,
                    aesthetics = c("colour", "fill"),
                    name = "CBG Count\nQuintile Breaks") +
  mapTheme() +
  labs(title = "How many different CBGs visit \neach restaurant (July)?")

grid.arrange(restaurants_cbg_jan, restaurants_cbg_jul, nrow = 1)
```

The following maps look at the number of CBGs per location for bars and concert/performance venues.
```{r echo=TRUE, message=FALSE, warning=FALSE}
#Bars
bars_cbg <-
  dat_cbg_visitors %>%
  filter(top_category == "Drinking Places (Alcoholic Beverages)") %>%
  group_by(safegraph_place_id) %>%
  summarize(Count = n()) %>%
  ggplot() +
  geom_sf(data = phl_cbg, fill = "grey40", color = "transparent") +
  geom_sf(aes(color = q5(Count)), size = 1) + 
  scale_fill_manual(values = palette5,
                    aesthetics = c("colour", "fill"),
                    name = "CBG Count\nQuintile Breaks") +
  mapTheme() +
  labs(title = "How many different CBGs visit bars?")

#Concert Venues
concert_cbg <-
  dat_cbg_visitors %>%
  filter(top_category == "Promoters of Performing Arts, Sports, and Similar Events" |
           top_category == "Performing Arts Companies") %>%
  group_by(safegraph_place_id) %>%
  summarize(Count = n()) %>%
  ggplot() +
  geom_sf(data = phl_cbg, fill = "grey40", color = "transparent") +
  geom_sf(aes(color = q5(Count)), size = 1) + 
  scale_fill_manual(values = palette5,
                    aesthetics = c("colour", "fill"),
                    name = "CBG Count\nQuintile Breaks") +
  mapTheme() +
  labs(title = "How many different CBGs visit \nconcert venues and performing arts venues?")

grid.arrange(bars_cbg, concert_cbg, nrow = 1)
```

### When are people travelling?

The following charts looks at the average visits by day in restaurants and other eating places. This shows the patterns for each of the subcategory under restaurant. 

We see an increase throughout the week until Friday and then decreased traffic over the weekend. This likely reflects the workforce that isn't coming into the city over the weekends.

```{r echo=TRUE, message=FALSE, warning=FALSE}
dat_day <- 
  dat2 %>% 
  select(safegraph_place_id, top_category, sub_category, popularity_by_day) %>%
  mutate(popularity_by_day = str_remove_all(popularity_by_day, pattern = "\\[|\\]")) %>%
  mutate(popularity_by_day = str_remove_all(popularity_by_day, pattern = "\\{|\\}")) %>%
  mutate(popularity_by_day = str_remove_all(popularity_by_day, pattern = '\\"|\\"')) %>% 
  mutate(popularity_by_day = str_split(popularity_by_day, pattern = ",")) %>%
  unnest(popularity_by_day) %>%
  separate(.,
           popularity_by_day,
           c("Day", "Visits"),
           sep = ":") %>%
  mutate(Visits = as.numeric(Visits))

week_order = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")

dat_day %>%
  filter(top_category == "Restaurants and Other Eating Places") %>%
  group_by(sub_category, Day) %>%
  summarize(Avg_Visits = mean(Visits)) %>%
  mutate(Day = factor(Day, levels = week_order)) %>% arrange(Day) %>% #order by day of the week
  ggplot(., aes(x = Day, y = Avg_Visits)) + 
  geom_col() +
  facet_wrap(~sub_category)
```

The same metric for bars (there are no sub-categories).
```{r echo=TRUE, message=FALSE, warning=FALSE}
dat_day %>%
  filter(top_category == "Drinking Places (Alcoholic Beverages)") %>%
  group_by(sub_category, Day) %>%
  summarize(Avg_Visits = mean(Visits)) %>%
  mutate(Day = factor(Day, levels = week_order)) %>% arrange(Day) %>% #order by day of the week
  ggplot(., aes(x = Day, y = Avg_Visits)) + 
  geom_col() +
  facet_wrap(~sub_category)

```

And for performing arts centers and companies.
```{r echo=TRUE, message=FALSE, warning=FALSE}
dat_day %>%
  filter(top_category == "Promoters of Performing Arts, Sports, and Similar Events" |
           top_category == "Performing Arts Companies") %>%
  group_by(top_category, Day) %>%
  summarize(Avg_Visits = mean(Visits)) %>%
  mutate(Day = factor(Day, levels = week_order)) %>% arrange(Day) %>% #order by day of the week
  ggplot(., aes(x = Day, y = Avg_Visits)) + 
  geom_col() +
  facet_wrap(~top_category)
```

Turning to traffic by hour of the day, the following chart shows foot average foot traffic in restaurants.  We have split them out by sub-category. At this level, it appears that most restaurant traffic does not occur at night, but rather in the middle of the day.
```{r echo=TRUE, message=FALSE, warning=FALSE}
dat_hour <- 
  dat2 %>% 
  select(safegraph_place_id, top_category, sub_category, popularity_by_hour, poi_cbg) %>%
  mutate(popularity_by_hour = str_remove_all(popularity_by_hour, pattern = "\\[|\\]")) %>%
  unnest(popularity_by_hour) %>%
  separate(.,
           popularity_by_hour,
           c("0", "1", "2", "3", "4", "5", "6", 
             "7", "8", "9", "10", "11", "12", 
             "13", "14", "15", "16", "17", "18",
             "19", "20", "21", "22", "23"),
           sep = ",") %>%
  pivot_longer(cols = 4:27,
               names_to = "Hour",
               values_to = "Count") %>%
  mutate(Hour = as.numeric(Hour),
         Count = as.numeric(Count))

#Restaurants
dat_hour %>%
  filter(top_category == "Restaurants and Other Eating Places") %>%
  group_by(sub_category, Hour) %>%
  summarize(Avg_Visits = mean(Count)) %>%
  ggplot(., aes(x = Hour, y = Avg_Visits)) + 
  geom_col() +
  facet_wrap(~sub_category)

```

We have also include the same metric for bars, which show a stronger pattern towards nighttime traffic. 
```{r echo=TRUE, message=FALSE, warning=FALSE}
#Bars
dat_hour %>%
  filter(top_category == "Drinking Places (Alcoholic Beverages)") %>%
  group_by(sub_category, Hour) %>%
  summarize(Avg_Visits = mean(Count)) %>%
  ggplot(., aes(x = Hour, y = Avg_Visits)) + 
  geom_col() +
  labs(title = "Average Traffic by Hour in Philadelphia Bars")
```

Same for concert venues and performing arts spaces, which have a high amount of traffic in the evening and early morning hours.
```{r echo=TRUE, message=FALSE, warning=FALSE}
#Concert Venues
dat_hour %>%
  filter(top_category == "Promoters of Performing Arts, Sports, and Similar Events" |
           top_category == "Performing Arts Companies") %>%
  group_by(top_category, Hour) %>%
  summarize(Avg_Visits = mean(Count)) %>%
  ggplot(., aes(x = Hour, y = Avg_Visits)) + 
  geom_col() +
  facet_wrap(~top_category)
```

### Number of visits/vistors. Hourly/weekly/monthly patterns? (Brian)
```{r}

#Need to modify creation of dat_hour above by addiong poi_cbg:
  #select(safegraph_place_id, top_category, sub_category, popularity_by_hour, poi_cbg) %>%

#Restaurant Analysis
#Filter Restaurants
dat_restaurants <-
dat_hour %>%
  filter(top_category == "Restaurants and Other Eating Places") 

#Plot map of restaurant locations
dat_restaurants %>%
  subset(Hour == 19 & Count < 2000) %>% 
  ggplot() + 
  geom_sf(data = phl_cbg, fill = "black", color = "transparent")+
  geom_sf(aes(colour = Count, geometry=geometry), alpha = .5, cex = .1) + 
  scale_colour_viridis(trans = "sqrt") +
  labs(title = "Restaurant Locations at 7pm")

#Merge CBGs with popularity by hour data
dat_restaurant_filter <-
  dat_restaurants %>%
  rename(., GEOID10 = poi_cbg) %>%
  group_by(GEOID10, Hour) %>%
  summarize(Avg_Popularity = mean(Count),
            Total_Visits = sum(Count)) %>%
  left_join(phl_cbg) %>% 
  st_as_sf() %>%
  mutate(Visits_Per_Area = Total_Visits / Shape__Area * 100)

#Plot of restaurant popularity between 7 and 8 pm
dat_restaurant_filter %>%
  subset(Hour == 19) %>% #selecting for one hour of the day
  ggplot() + 
  geom_sf(data = phl_cbg, fill = "#440255", color = "transparent") +
  geom_sf(aes(fill = Visits_Per_Area), color = "transparent") + 
  scale_fill_viridis(trans = "sqrt") +
  labs(title = "Total Restaurant Visits by CBG") 

#Animation of restaurant popularity by hour
restaurant.animation.data <-
    dat_restaurant_filter %>%
    st_sf() %>%
    mutate(Pop_String = case_when(Visits_Per_Area < .4 ~ "5",
                              Visits_Per_Area >= .4 & Visits_Per_Area <.8 ~ "4",
                              Visits_Per_Area >= .8 & Visits_Per_Area <1.2 ~ "3",
                              Visits_Per_Area >= 1.2 & Visits_Per_Area <1.6 ~ "2",
                              Visits_Per_Area >= 2 ~ "1")) %>%
    mutate(Pop_String  = fct_relevel(Pop_String, "5","4","3","2","1"))

restaurant_animation <-
  ggplot() +
  geom_sf(data = phl_cbg, fill = "#440255", color = "transparent") +   
  geom_sf(data = restaurant.animation.data, aes(fill = Pop_String), color = "transparent") +
    scale_fill_manual(values = palette5) +
    labs(title = "Restaurant Popularity by Hour",
         subtitle = "One Hour Intervals: {current_frame}") +
  theme(panel.background = element_rect(fill = "black"),
         panel.grid.major = element_line(color = "transparent"),
          panel.grid.minor = element_line(colour = "transparent")) +
    transition_manual(Hour)
  

animate(restaurant_animation, duration=20, renderer = gifski_renderer())

#Bar Analysis
#Filter Bars
dat_bars <-
dat_hour %>%
  filter(top_category == "Drinking Places (Alcoholic Beverages)") 

#Plot map of bar locations
dat_bars %>%
  subset(Hour == 23 & Count < 2400) %>% 
  ggplot() + 
  geom_sf(data = phl_cbg, fill = "black", color = "transparent")+
  geom_sf(aes(colour = Count, geometry=geometry), alpha = .5, cex = .2) + 
  scale_colour_viridis(trans = "sqrt") +
  labs(title = "Bar Locations and Popularity at 11pm")

#Merge CBGs with popularity by hour data
dat_bars_filter <-
  dat_bars %>%
  rename(., GEOID10 = poi_cbg) %>%
  group_by(GEOID10, Hour) %>%
  summarize(Avg_Popularity = mean(Count),
            Total_Visits = sum(Count)) %>%
  left_join(phl_cbg) %>% 
  st_as_sf() %>%
  mutate(Visits_Per_Area = Total_Visits / Shape__Area * 100)

dat_bars_filter$Visits_Per_Area[is.na(dat_bars_filter$Visits_Per_Area)] <- 0

#Plot of bar popularity between 11pm and 12 am
dat_bars_filter %>%
  subset(Hour == 23) %>% #selecting for one hour of the day
  ggplot() + 
  geom_sf(data = phl_cbg, fill = "#440255", color = "transparent") +
  geom_sf(aes(fill = Visits_Per_Area, geometry=geometry), color = "transparent") + 
  scale_fill_viridis(trans = "sqrt") +
  labs(title = "Total Bar Visits by CBG, 11pm-12am") 

#Animation of bar popularity by hour
bar.animation.data <-
    dat_bars_filter %>%
    st_sf() %>%
    mutate(Pop_String = case_when(Visits_Per_Area == 0 ~ "5",
                              Visits_Per_Area > 0 & Visits_Per_Area <.2 ~ "4",
                              Visits_Per_Area >= .2 & Visits_Per_Area <.4 ~ "3",
                              Visits_Per_Area >= .4 & Visits_Per_Area <.6 ~ "2",
                              Visits_Per_Area >= .6 ~ "1")) %>%
    mutate(Pop_String  = fct_relevel(Pop_String, "5","4","3","2","1"))

bar_animation <-
  ggplot() +
  geom_sf(data = phl_cbg, fill = "#440255", color = "transparent", ) +  
  geom_sf(data = bar.animation.data, aes(fill = Pop_String), color = "transparent") +
    scale_fill_manual(values = palette5) +
    labs(title = "Bar Popularity by Hour",
         subtitle = "One Hour Intervals: {current_frame}") +
  theme(panel.background = element_rect(fill = "black"),
         panel.grid.major = element_line(color = "transparent"),
          panel.grid.minor = element_line(colour = "transparent")) +
    transition_manual(Hour)
  

animate(bar_animation, duration=20, renderer = gifski_renderer())


```

### Where does nightlife happen?

```{r}
#Census Block Groups
dat_cbg2 <-
  dat %>%
  select(safegraph_place_id,
         poi_cbg, 
         raw_visit_counts, 
         raw_visitor_counts, 
         median_dwell, 
         distance_from_home)%>%
  rename(., GEOID10 = poi_cbg) %>%
  group_by(GEOID10) %>%
  summarize(Avg_Visits = mean(raw_visit_counts),
            Avg_Visitors = mean(raw_visitor_counts),
            Total_Visits = sum(raw_visit_counts),
            Total_Visitors = sum(raw_visitor_counts),
            Avg_Dwell = mean(median_dwell),
            Avg_DistHome = mean(distance_from_home)) %>%
  left_join(phl_cbg) %>% 
  st_as_sf() %>%
  mutate(Visits_Per_Area = Total_Visits / Shape__Area) %>%
  mutate(Visitors_Per_Area = Total_Visitors / Shape__Area)

dat_cbg2 %>%
  subset(Total_Visits < 1900000) %>% #removing outliers to see patterns in the data
  ggplot() + 
  geom_sf(data = phl_cbg, fill = "grey40", color = "transparent") +
  geom_sf(aes(fill = Total_Visits), color = "transparent") + 
  scale_fill_viridis() +
  labs(title = "Total SafeGraph Visits by CBG")

dat_cbg2 %>%
  subset(Total_Visitors < 1900000) %>% #removing outliers to see patterns in the data
  ggplot() + 
  geom_sf(data = phl_cbg, fill = "grey40", color = "transparent") +
  geom_sf(aes(fill = Total_Visits), color = "transparent") + 
  scale_fill_viridis() +
  labs(title = "Total SafeGraph Visitors by CBG")

dat_cbg2 %>%
  #subset(Avg_Visits < 1500) %>% #removing outliers to see patterns in the data
  ggplot() + 
  geom_sf(data = phl_cbg, fill = "grey40", color = "transparent") +
  geom_sf(aes(fill = Visits_Per_Area), color = "transparent") + 
  scale_fill_viridis() +
  labs(title = "Total SafeGraph Visits by CBG, Adjusted by CBG Area")

dat_cbg2 %>%
  #subset(Avg_Visits < 1500) %>% #removing outliers to see patterns in the data
  ggplot() + 
  geom_sf(data = phl_cbg, fill = "grey40", color = "transparent") +
  geom_sf(aes(fill = Visitors_Per_Area), color = "transparent") + 
  scale_fill_viridis() +
  labs(title = "Total SafeGraph Visitors by CBG, Adjusted by CBG Area")

```

### What are the major nightlife corridors? (Brian)
```{r}
#Creating a fishnet
phillyBoundary <- phl_zip %>%
  select(geometry) %>%
  st_union() %>%
   st_transform('ESRI:102271') %>%
  st_as_sf() 

fishnet <- 
  st_make_grid(phillyBoundary, cellsize = 500, square = FALSE) %>%
  st_sf() %>%
  mutate(uniqueID = rownames(.))

#Plot fishnet
ggplot() + 
  geom_sf(data = fishnet, fill = "#440255", color = "black")

#Transform restaurants coordinates (Takes a long time)
dat_restaurants_grid <- dat_restaurants %>%
  select(geometry) %>%
  na.omit() %>%
  st_as_sf() %>%
  st_transform('ESRI:102271') %>% 
    distinct()

#Merge restaurants with fishnet  
restaurant_net <- 
  dplyr::select(dat_restaurants_grid) %>% 
  mutate(countRestaurant = 1) %>% 
  aggregate(., fishnet, sum) %>%
  mutate(countRestaurant = replace_na(countRestaurant, 0),
         uniqueID = rownames(.),
         cvID = sample(round(nrow(fishnet) / 24), size=nrow(fishnet), replace = TRUE))

#Plot restaurant fishnet
ggplot() +
  geom_sf(data = restaurant_net, aes(fill = countRestaurant)) +
  scale_fill_viridis(trans = "sqrt") +
  labs(title = "Count of Restaurants per Grid Space") +
  mapTheme()

#Transform bar coordinates (Takes a long time)
dat_bars_grid <- dat_bars %>%
  select(geometry) %>%
  #na.omit() %>%
  st_as_sf() %>%
  st_transform('ESRI:102271') %>% 
    distinct()

#Merge bars with fishnet  
bar_net <- 
  dplyr::select(dat_bars_grid) %>% 
  mutate(countBar = 1) %>% 
  aggregate(., fishnet, sum) %>%
  mutate(countBar = replace_na(countBar, 0),
         uniqueID = rownames(.),
         cvID = sample(round(nrow(fishnet) / 24), size=nrow(fishnet), replace = TRUE))

#Plot bar fishnet
ggplot() +
  geom_sf(data = bar_net, aes(fill = countBar)) +
  scale_fill_viridis(trans = "sqrt") +
  labs(title = "Count of Bars per Grid Space") +
  mapTheme()

```


###What is the relationship to transit corridors? (Brian)
```{r}
library(tidyverse)
library(tidycensus)
library(sf)
library(kableExtra)

#Septa stops (Brian)
septaStops <- 
  rbind(
    st_read("https://opendata.arcgis.com/datasets/8c6e2575c8ad46eb887e6bb35825e1a6_0.geojson") %>% 
      mutate(Line = "El") %>%
      select(Station, Line),
    st_read("https://opendata.arcgis.com/datasets/2e9037fd5bef406488ffe5bb67d21312_0.geojson") %>%
      mutate(Line ="Broad_St") %>%
      select(Station, Line)) %>%
  st_transform(st_crs(phila))  


#Plot location of restaurants and Septa Stations
ggplot() +
  geom_sf(data = restaurant_net, aes(fill = countRestaurant)) +
  scale_fill_viridis(trans = "sqrt") +
  geom_sf(data=septaStops, 
          aes(colour = Line), 
          show.legend = "point", cex = 1) +
  scale_colour_manual(values = c("orange","white")) +
  labs(title = "Count of Restaurants and Septa Stations") +
  mapTheme()

#Plot location of bars and Septa Stations
ggplot() +
  geom_sf(data = bar_net, aes(fill = countBar)) +
  scale_fill_viridis(trans = "sqrt") +
  geom_sf(data=septaStops, 
          aes(colour = Line), 
          show.legend = "point", cex = 1) +
  scale_colour_manual(values = c("orange","white")) +
  labs(title = "Count of Bars and Septa Stations") +
  mapTheme()

```

### How long do people spend in establishments?

[TBD]