---
title: "MUSA Practicum Nighttime Economy: Exploratory Analysis"
author: "Maddy Kornhauser, Brian Rawn, Sabrina Lee"
date: "3/2/2021"
output:
  github_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	width=6000,
	out.width = "100%"
)
```

# 0. Weekly Updates 3/2/2021

This week, we have refined and explained our list of corridor metrics, explored potential model features for our prediction use case, and begun wireframing. A detailed update can be found below:

* Corridor metrics: 
  + Corridor popularity over time: This metric looks at the volume of trips over time. Since we want to stay away from generating raw counts, we will likely show this as a percent change in foot traffic over time. We will likely use the popularity_by_hour and popularity_by_day variables in the
SafeGraph dataset. 
  + Demographic characteristics: Based on the origin census tracts, we will look at the following demographic indicators. These will likely be presented as a weighted average based on how many visitors from that census tract: Age (Median Age variable), Income (Median Household Income variable), Race / Ethnicity, Educational Attainment, Household composition. (**Question:  What is the best way to represent categorical variables here?**)
  + Average distance travelled to nightlife destinations: the weighted average of the origin block group centroid and the desitnation corridor centroid.
  + Average dwell time by commercial use: Aggregate SafeGraph's median_dwell variable by businesses in the commercial corridor.
  + The percentage drop in trips that nightlife establishments experienced as a result of COVID-19.
* Prediction Parameters: We plan to predict the number of trips by commercial corridor. This prediciton will be transformed into a percentage increase or decrease relative to average traffic.
* Feature Engineering:
  + So far distance_from_home appears to have the strongest relationship with the number of trips.
  + We have started looking at clustering and counts of amenities along corridors, but there doesn't seem to be a particularly strong relationship. (**Question:  Should we be using SafeGraph for the creation of these "amentity" variables or something more comprehensive, like OSM?**)
  + Tips on how to mine SafeGraph data for meaningful variables.
  + Want to incorporate outside data next: Census, transit, land use, vacancy, building foot print.
* Wireframing:  We created brief wireframes for popularity by types of nightlife/neighborhood, and where people are coming from. Next, we want to work on creating wireframes for visualizing change pre- and post-COVID as well as visualizing differences by visitor demographic.

# 1. Use Case Development

## Purpose

Our goal is to help business improvement districts (BIDs) and small businesses understand the patterns of nightlife across Philadelphia’s commercial corridors in order to further recovery efforts following the COVID-19 pandemic. 

## The Tool

An interactive data dashboard that allows users to understand nightlife patterns at the corridor level across Philadelphia. This dashboard would, for each corridor, answer the questions of “who, what, where, and when?” of nightlife visitors. The dashboard would also allow for comparisons to other corridors. Potential metrics that would be displayed include:

* Corridor popularity (volume of trips) over time. 
* The percentage of visitors attributable to different origin census tracts and a breakdown of associated demographic characteristics.
* Average distance traveled to restaurants/bars/theaters.
* Average dwell time by commercial use.
* The percentage drop in trips that restaurants/bars/theaters experienced as a result of COVID-19.
* For each, a comparison to the average of other commercial corridors.

**(In progress)** In addition, the tool will have the additional functionality of being able to predict the future popularity of a commercial corridor given changes in time (hour, day, month), weather, volume of commercial establishments (count or floor area), and retail mix. This predictive functionality would help inform decisions relating to the regulation and expansion of nightlife activities.

## Applications

How will this tool be utilized for economic development? Potential applications include use by BIDs and small businesses to:

* Understand peak trip times by use and location in order to inform parking, transit, or commercial policies.
* Understand relative popularity of commercial corridors to inform decisions to grant licenses or small business assistance.
* Understand origins and demographics of visitors to more effectively target marketing resources. 
* Understand the future effect of more commercial establishments or square footage in a given commercial corridor.

# 2. The SafeGraph Dataset

Safegraph uses anonymized cell phone GPS data to record trips to commercial points of interest. This data can tell us from where a trip was made, what time the trip was made, and how long the individual stated at the point of interest.

* Pros: 
  + Brand new dataset. 
  + Lots of unexplored applications and potential insights

* Cons: 
  + Data is new and has a significant amount of incorrectly attributed trips, especially in urban areas

# 3. Exploratory Data Analysis

```{r load packages, include=FALSE}
library(tidyverse)
library(tidycensus)
library(sf)
library(lubridate)
library(datetime)
library(viridis)
library(gridExtra)
library(kableExtra)
library(dplyr)
library(ggplot2)
library(gganimate)
library(riem)

census_api_key("30bae866445f7bea5ea57b8f5fba60b11b8b145e", overwrite = TRUE, install = TRUE) 

qBr <- function(df, variable, rnd) {
  if (missing(rnd)) {
    as.character(quantile(round(df[[variable]],0),
                          c(.01,.2,.4,.6,.8), na.rm=T))
  } else if (rnd == FALSE | rnd == F) {
    as.character(formatC(quantile(df[[variable]]), digits = 3),
                 c(.01,.2,.4,.6,.8), na.rm=T)
  }
}
q5 <- function(variable) {as.factor(ntile(variable, 5))}
palette3 <- viridis_pal()(3)
palette5 <- viridis_pal()(5)
mapTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 14,colour = "black"),
    plot.subtitle=element_text(face="italic"),
    plot.caption=element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),axis.title = element_blank(),
    axis.text = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=2)
  )
}

plotTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 16,colour = "black"),
    plot.subtitle = element_text(face="italic"),
    plot.caption = element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    panel.grid.major = element_line("grey80", size = 0.1),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=2),
    strip.background = element_rect(fill = "grey80", color = "white"),
    strip.text = element_text(size=12),
    axis.title = element_text(size=12),
    axis.text = element_text(size=10),
    plot.background = element_blank(),
    legend.background = element_blank(),
    legend.title = element_text(colour = "black", face = "italic"),
    legend.text = element_text(colour = "black", face = "italic"),
    strip.text.x = element_text(size = 14)
  )
}

setwd("~/GitHub/musa_practicum_nighttime")
```

```{r load & wrangle data}
dat <- read.csv("./data/moves_2018.csv")
phila <- st_read("./demo/phila.geojson", quiet = TRUE)

dat2 <- dat %>% 
  dplyr::select(safegraph_place_id, 
                date_range_start, 
                date_range_end, 
                raw_visit_counts,
                raw_visitor_counts, 
                visits_by_day, 
                poi_cbg, 
                visitor_home_cbgs, 
                visitor_daytime_cbgs, 
                visitor_work_cbgs, 
                visitor_country_of_origin,
                distance_from_home, 
                median_dwell, 
                bucketed_dwell_times, 
                related_same_day_brand, 
                related_same_month_brand, 
                popularity_by_hour, 
                popularity_by_day, 
                device_type) %>%
  left_join(., phila, by = "safegraph_place_id") %>% 
  st_as_sf() %>%
  st_transform('ESRI:102728')

#Block group shapefiles (projected & unprojected)
phl_cbg <- 
  st_read("http://data.phl.opendata.arcgis.com/datasets/2f982bada233478ea0100528227febce_0.geojson", quiet = TRUE) %>%
  mutate(GEOID10 = as.numeric(GEOID10))%>%
  st_transform('ESRI:102728') 
phl_cbg_unproj <- 
  st_read("http://data.phl.opendata.arcgis.com/datasets/2f982bada233478ea0100528227febce_0.geojson", quiet = TRUE) %>%
  mutate(GEOID10 = as.numeric(GEOID10),
         Lat = as.numeric(INTPTLAT10),
         Lon = as.numeric(INTPTLON10))

#Boundaries shapefiles (projected & unprojected)
PHL_boundary <- 
  st_read("https://opendata.arcgis.com/datasets/405ec3da942d4e20869d4e1449a2be48_0.geojson", quiet = TRUE)%>%
  st_transform('ESRI:102728') 
PHL_boundary_unproj <- 
  st_read("https://opendata.arcgis.com/datasets/405ec3da942d4e20869d4e1449a2be48_0.geojson", quiet = TRUE)

#Corridor shapefiles (projected & unprojected)
phl_corridors <- st_read("http://data.phl.opendata.arcgis.com/datasets/f43e5f92d34e41249e7a11f269792d11_0.geojson", quiet = TRUE)%>%
  st_transform('ESRI:102728') 
phl_corridors_unproj <- st_read("http://data.phl.opendata.arcgis.com/datasets/f43e5f92d34e41249e7a11f269792d11_0.geojson", quiet = TRUE)

#Neighborhood shapefiles (projected & unprojected)
phl_nhoods <- 
  st_read("https://raw.githubusercontent.com/azavea/geo-data/master/Neighborhoods_Philadelphia/Neighborhoods_Philadelphia.geojson", 
          quiet = TRUE) %>%
  st_transform('ESRI:102728')
phl_nhoods_unproj <- 
  st_read("https://raw.githubusercontent.com/azavea/geo-data/master/Neighborhoods_Philadelphia/Neighborhoods_Philadelphia.geojson", 
          quiet = TRUE)

#Planning district shapefiles (projected & unprojected)
phl_dist <-
  st_read("http://data.phl.opendata.arcgis.com/datasets/0960ea0f38f44146bb562f2b212075aa_0.geojson", 
          quiet = TRUE) %>%
  st_transform('ESRI:102728')

phl_dist_unproj <-
  st_read("http://data.phl.opendata.arcgis.com/datasets/0960ea0f38f44146bb562f2b212075aa_0.geojson", 
          quiet = TRUE)

#Philadelphia zip codes
phl_zip <- st_read("http://data.phl.opendata.arcgis.com/datasets/b54ec5210cee41c3a884c9086f7af1be_0.geojson", quiet = TRUE) %>%
  st_transform('ESRI:102728') %>%
  mutate(CODE = as.numeric(CODE))
```

## Philadelphia nightlife establishments

Our first research question is where Philadelphia nightlife establishments are located across the city. The following maps indicate where businesses that contribute to the city's nightlife economy are located. The categories include:

* Bars (Drinking Places)
* Restaurants 
* Hotels (Travel Accomodation)
* Casinos (Gambling Industries)
* Arts Venues (Promotes of Performing Arts)

Figure 3.1 below shows the spatial patterns of the business categories. Bars and restaurants represent the highest number of businesses which are spread across the city. Hotels are mostly clustered in the central district of the city and near the airport in the southwest portion of the city. There are far fewer casinos and arts venues.

Throughout the analysis, we pay special attention to bars and restaurants, as these organizations are well distributed throughout the city and apply to a local Philadelphia customer base.

```{r establishment locations}
dat2 %>%
  filter(top_category == "Drinking Places (Alcoholic Beverages)" |
           top_category == "Restaurants and Other Eating Places" |
           # top_category == "Traveler Accommodation" |
           # top_category == "Gambling Industries" |
           top_category == "Promoters of Performing Arts, Sports, and Similar Events" |
           top_category == "Performing Arts Companies") %>%
  ggplot() + 
  geom_sf(data = phl_cbg, fill = "grey80", color = "transparent") +
  geom_sf(color = "red", size = .1) +
  labs(title = "Location of Nightlife Establishments",
       subtitle = "Figure X.X") +
  facet_wrap(~top_category, nrow = 1) +
  mapTheme()
```

To look at the spatial patterns another way, we review the distribution of businesses  with a fishnet grid. The fishnet allows us to visualize clusters and hotspots. Starting first with restaurants, Figure X.X below demonstrates the spatial patterns of these businesses.  Though restaurants are well distributed throughout the city, there appears to be a higher concentration of yellow cells (indicating a higher count of restaurants) in the central part of the city. This corresponds to what we observe in the point data analysis. 
```{r unnesting hour dataset}
#Unnesting popularity by hour variable
dat_hour <- 
  dat2 %>% 
  select(safegraph_place_id, top_category, sub_category, popularity_by_hour, poi_cbg, median_dwell) %>%
  mutate(popularity_by_hour = str_remove_all(popularity_by_hour, pattern = "\\[|\\]")) %>%
  unnest(popularity_by_hour) %>%
  separate(.,
           popularity_by_hour,
           c("0", "1", "2", "3", "4", "5", "6", 
             "7", "8", "9", "10", "11", "12", 
             "13", "14", "15", "16", "17", "18",
             "19", "20", "21", "22", "23"),
           sep = ",") %>%
  pivot_longer(cols = 4:27,
               names_to = "Hour",
               values_to = "Count") %>%
  mutate(Hour = as.numeric(Hour),
         Count = as.numeric(Count))

#Creating a fishnet
phillyBoundary <- 
  phl_zip %>%
  select(geometry) %>%
  st_union() %>%
   st_transform('ESRI:102728') %>% 
  st_as_sf() 

fishnet <- 
  st_make_grid(phillyBoundary, cellsize = 1500, square = FALSE) %>%
  .[phillyBoundary] %>% #MK added this line
  st_sf() %>%
  mutate(uniqueID = rownames(.))

# #Plot fishnet
# ggplot() +
#   geom_sf(data = fishnet, fill = "#440255", color = "black")

```

Figure X.X below includes other sectors that contribute to Philadelphia's nightlife economy such as hotels and casinos. This figure combines the fishnets into a single panel that allows us to compare distribution of industries across business type.
```{r}
#Filter restaurants
restaurants <- dat2 %>%
  filter(top_category == "Restaurants and Other Eating Places")%>%
    st_transform('ESRI:102728')%>%
  mutate(Legend = "Restaurants")

#aggregate restaurant count by fishnet cell
restaurants_net <-
  dplyr::select(restaurants) %>% 
  mutate(countRestaurants = 1) %>% 
  aggregate(., fishnet, sum) %>%
  mutate(countRestaurants = replace_na(countRestaurants, 0),
         uniqueID = rownames(.),
         cvID = sample(round(nrow(fishnet) / 24), 
                       size=nrow(fishnet), replace = TRUE))

#Bars
bars <- dat2 %>%
  filter(top_category == "Drinking Places (Alcoholic Beverages)")%>%
    st_transform('ESRI:102728')%>%
  mutate(Legend = "Bars")

#aggregate bars by fishnet cell
bars_net <-
  dplyr::select(bars) %>% 
  mutate(countBars = 1) %>% 
  aggregate(., fishnet, sum) %>%
  mutate(countBars = replace_na(countBars, 0),
         uniqueID = rownames(.),
         cvID = sample(round(nrow(fishnet) / 24), 
                       size=nrow(fishnet), replace = TRUE)) %>%
  mutate(Legend = "Bars")

# #Casinos
# casinos <- dat2 %>%
#   filter(top_category == "Gambling Industries") %>%
#   st_transform('ESRI:102728') %>%
#   mutate(Legend = "Casinos")
# 
# casinos_net <-
#   dplyr::select(casinos) %>% 
#   mutate(countCasinos = 1) %>% 
#   aggregate(., fishnet, sum) %>%
#   mutate(countCasinos = replace_na(countCasinos, 0),
#          uniqueID = rownames(.),
#          cvID = sample(round(nrow(fishnet) / 24), 
#                        size=nrow(fishnet), replace = TRUE))

#Performing arts
performingarts <- dat2 %>%
  filter(top_category == "Promoters of Performing Arts, Sports, and Similar Events" |
           top_category == "Performing Arts Companies") %>%
  st_transform('ESRI:102728') %>%
  mutate(Legend = "Performing Arts")

performingarts_net <-
  dplyr::select(performingarts) %>% 
  mutate(countPerformingarts = 1) %>% 
  aggregate(., fishnet, sum) %>%
  mutate(countPerformingarts = replace_na(countPerformingarts, 0),
         uniqueID = rownames(.),
         cvID = sample(round(nrow(fishnet) / 24), 
                       size=nrow(fishnet), replace = TRUE))

# #Hotels
# hotels <- dat2 %>%
#   filter(top_category == "Traveler Accommodation") %>%
#   st_transform('ESRI:102728') %>%
#   mutate(Legend = "Hotels")
# 
# hotels_net <-
#   dplyr::select(hotels) %>% 
#   mutate(countHotels = 1) %>% 
#   aggregate(., fishnet, sum) %>%
#   mutate(countHotels = replace_na(countHotels, 0),
#          uniqueID = rownames(.),
#          cvID = sample(round(nrow(fishnet) / 24), 
#                        size=nrow(fishnet), replace = TRUE))

# Combining fishnets into a single dataframe
vars_net <- 
  rbind(restaurants, 
        bars, 
        performingarts
        # , 
        # casinos, 
        # hotels
        ) %>%
  st_join(., fishnet, join=st_within) %>%
  st_drop_geometry() %>%
  group_by(uniqueID, Legend) %>%
  dplyr::summarize(count = n()) %>%
    full_join(fishnet) %>%
    spread(Legend,count, fill=0) %>%
    st_sf() %>%
    dplyr::select(-`<NA>`) %>%
    na.omit() %>%
    ungroup()

vars_net.long <- 
  gather(vars_net, Variable, value, -geometry, -uniqueID)

vars <- unique(vars_net.long$Variable)
mapList <- list()

#Plotting small multiple maps
for(i in vars){
  mapList[[i]] <- 
    ggplot() +
      geom_sf(data = filter(vars_net.long, Variable == i), aes(fill=value), colour=NA) +
      scale_fill_viridis(name="") +
      labs(title=i) +
      mapTheme()}

do.call(grid.arrange, c(mapList, ncol=3, top="Count of Nightlife Businesses per Fishnet", bottom = "Figure X.X"))
```

## Nightlife hours

Next, we explore when visitors make trips to nightlife establishments. To do this, we worked with the popularity_by_hour SafeGraph variable, which sums the total number of visitors by hour for each month.

Figure X.X shows the average foot traffic by business type over the course of a day. This graphic indicates that though certain business types are considererd part of the nightlife economy, they do not exclusively experience traffic in the evenings. Restaurants are a good example of this, where the data indicates that the highest levels of traffic occur in the middle of the day. Arts venues, on the other hand, have a clear patterns indicating that they are more popular later in the day.
```{r traffic by nightlife establishment}
dat_hour %>%
  filter(top_category == "Drinking Places (Alcoholic Beverages)" |
           top_category == "Restaurants and Other Eating Places" |
           # top_category == "Traveler Accommodation" |
           # top_category == "Gambling Industries" |
           top_category == "Promoters of Performing Arts, Sports, and Similar Events" |
           top_category == "Performing Arts Companies") %>%
  group_by(Hour, top_category) %>%
  dplyr::summarize(Count = mean(Count)) %>%
  ggplot(., aes(x = Hour, y = Count)) + 
  geom_col() +
  labs(title = "Philadelphia Nightlife Organizations, Average Traffic by Hour",
       subtitle = "Figure X.X") +
  facet_wrap(~top_category, scales = "free") +
  plotTheme()
```

The following animation shows restaurant visitor counts by census block group over the course of the day. We observe the highest amount of traffic in center city in the middle of the day and then a resurgence of traffic elsewhere in the city in the evening. This suggests that people dine at neighborhood establishments close to their home in the evening.
```{r}
dat_restaurants <-
  dat_hour %>%
  filter(top_category == "Restaurants and Other Eating Places") 

dat_restaurant_filter <-
  dat_restaurants %>%
  dplyr::rename(., GEOID10 = poi_cbg) %>%
  dplyr::group_by(GEOID10, Hour) %>%
  dplyr::summarize(Avg_Popularity = mean(Count),
            Total_Visits = sum(Count)) %>%
  left_join(phl_cbg) %>% 
  st_as_sf() %>%
  mutate(Visits_Per_Area = Total_Visits / Shape__Area * 100)

#Animation of restaurant popularity by hour
restaurant.animation.data <-
    dat_restaurant_filter %>%
    st_sf() %>%
    mutate(Pop_String = case_when(Visits_Per_Area < .4 ~ "5",
                              Visits_Per_Area >= .4 & Visits_Per_Area <.8 ~ "4",
                              Visits_Per_Area >= .8 & Visits_Per_Area <1.2 ~ "3",
                              Visits_Per_Area >= 1.2 & Visits_Per_Area <1.6 ~ "2",
                              Visits_Per_Area >= 2 ~ "1")) %>%
    mutate(Pop_String  = fct_relevel(Pop_String, "5","4","3","2","1"))

restaurant_animation <-
  ggplot() +
  geom_sf(data = phl_cbg, fill = "#440255", color = "transparent") +   
  geom_sf(data = restaurant.animation.data, aes(fill = Pop_String), color = "transparent") +
    scale_fill_manual(values = palette5) +
    labs(title = "Restaurant Popularity by Hour",
         subtitle = "One Hour Intervals: {current_frame}") +
  theme(panel.background = element_rect(fill = "black"),
         panel.grid.major = element_line(color = "transparent"),
          panel.grid.minor = element_line(colour = "transparent")) +
    transition_manual(Hour)
  

animate(restaurant_animation, duration=20, renderer = gifski_renderer())
```

The following animation shows the same metric for bars. We observe traffic increasing throughout the day starting in the afternoon.
```{r bar animation}
#Bar Analysis
#Filter Bars
dat_bars <- dat_hour %>%
  filter(top_category == "Drinking Places (Alcoholic Beverages)") 

#Merge CBGs with popularity by hour data
dat_bars_filter <-
  dat_bars %>%
  dplyr::rename(., GEOID10 = poi_cbg) %>%
  dplyr::group_by(GEOID10, Hour) %>%
  dplyr::summarize(Avg_Popularity = mean(Count),
            Total_Visits = sum(Count)) %>%
  left_join(phl_cbg) %>% 
  st_as_sf() %>%
  mutate(Visits_Per_Area = Total_Visits / Shape__Area * 100)

#Animation of bar popularity by hour
bar.animation.data <-
    dat_bars_filter %>%
    st_sf() %>%
    mutate(Pop_String = case_when(Visits_Per_Area == 0 ~ "5",
                              Visits_Per_Area > 0 & Visits_Per_Area <.2 ~ "4",
                              Visits_Per_Area >= .2 & Visits_Per_Area <.4 ~ "3",
                              Visits_Per_Area >= .4 & Visits_Per_Area <.6 ~ "2",
                              Visits_Per_Area >= .6 ~ "1")) %>%
    mutate(Pop_String  = fct_relevel(Pop_String, "5","4","3","2","1"))

bar_animation <-
  ggplot() +
  geom_sf(data = phl_cbg, fill = "#440255", color = "transparent", ) +
  geom_sf(data = bar.animation.data, aes(fill = Pop_String), color = "transparent") +
    scale_fill_manual(values = palette5) +
    labs(title = "Bar Popularity by Hour",
         subtitle = "One Hour Intervals: {current_frame}") +
  theme(panel.background = element_rect(fill = "black"),
         panel.grid.major = element_line(color = "transparent"),
          panel.grid.minor = element_line(colour = "transparent")) +
    transition_manual(Hour)

animate(bar_animation, duration=20, renderer = gifski_renderer())
```

## Trip Flows

### Origins & destinations

The SafeGraph data fundamentally captures flow of people across space by connecting a series of origins and destinations. By mapping the origins and destinations of trips taking place across Philadlephia, we can observe which regions of the city draw from a larger crowd across the city, and which areas cater to a more local population.

The following maps look at the origins and destinations for trips to restaurants, bars, and arts venues across the city. Specifically, it shows the distance between the centroid of the origin neighborhood to the centroid of the destination commercial corridor. We selected 6 corridors across Philadelphia in different areas of the city that we felt represented a diverse array of commercial corridors in the city. Going forward, we would like to run a K-means clustering test to help sort the corridors into like categories that will help us understand the distinct profiles of corridors in Philadelphia.

The code draws each line segments based on lat/long coordinates of the origin an destination centroid (could not get this to work with projected data). Note that trips to destinations located outside of the commercial corridors are left off of these maps. 

The following code block loads unprojected shapefiles for the block groups and city boundary as well as wrangle the data into a list of individual origins and destinations for nightlife-related businesses.
```{r flows data wrangling}
flows <- dat2 %>%
  filter(top_category == "Drinking Places (Alcoholic Beverages)" |
           top_category == "Restaurants and Other Eating Places" |
           top_category == "Promoters of Performing Arts, Sports, and Similar Events" |
           top_category == "Performing Arts Companies") %>% #filter for business types
  st_join(phl_corridors) %>% #join destinations to phl corridors (apply corridor destination to each trip)
  st_as_sf() %>%
  st_drop_geometry() %>% 
  select(safegraph_place_id, 
         date_range_start,
         top_category,
         poi_cbg,
         visitor_home_cbgs,
         NAME.y) %>% #select columns
  mutate(visitor_home_cbgs = str_remove_all(visitor_home_cbgs, pattern = "\\[|\\]")) %>%
  mutate(visitor_home_cbgs = str_remove_all(visitor_home_cbgs, pattern = "\\{|\\}")) %>%
  mutate(visitor_home_cbgs = str_remove_all(visitor_home_cbgs, pattern = '\\"|\\"')) %>%
  mutate(visitor_home_cbgs = str_split(visitor_home_cbgs, pattern = ",")) %>%
  unnest(visitor_home_cbgs) %>% #unnest visitor cbg column
  separate(.,
           visitor_home_cbgs,
           c("visitor_cbg", "count"),
           sep = ":") %>% #separate count column from visitor cbg
  mutate(count = as.numeric(count),
         visitor_cbg = as.numeric(visitor_cbg)) %>%
  dplyr::rename(., corridor_dest = NAME.y) %>%
  drop_na(corridor_dest) #Remove trips with destinations outside of corridors

#prepare PHL boundary file to coordinates for plotting
PHL_boundary_unproj <- 
  PHL_boundary_unproj %>% 
  st_coordinates() # split out coordinates 

PHL_boundary_unproj <-
  as.data.frame(PHL_boundary_unproj) # save as dataframe

#----MAP V1: NEIGHBORHOOD CENTROID TO CORRIDOR CENTROIDS
#Generate CBG centroid
phl_cbg_cent <- 
  phl_cbg_unproj %>% 
  select(GEOID10, geometry) %>% 
  st_centroid()

#Generate neighborhood centroids
phl_nhood_cent <- 
  phl_nhoods_unproj %>% 
  select(name) %>%
  st_centroid()

#Generate corridor centroid
phl_corr_cent <- 
  phl_corridors_unproj %>% 
  select(NAME, geometry) %>% 
  st_centroid()

#Join CBG centroid to neighborhood shapefile
phl_cbg_nhood <- 
  st_join(phl_nhoods_unproj, phl_cbg_cent, join = st_intersects) %>%
  select(GEOID10, name) %>%
  st_centroid()

flows_nhood <- flows %>%
  left_join(phl_cbg_nhood, by=c("visitor_cbg"="GEOID10")) %>% #join visitor CBGs to nhoods
  dplyr::rename(., nhood_origin = name) %>% #cleanup columns for clarity
  drop_na(nhood_origin) %>% #dropping trips outside of Philadelphia
  st_as_sf() %>%
  st_drop_geometry() %>% 
  group_by(nhood_origin, top_category, corridor_dest) %>% #grouping trip counts by origin neighborhood
  summarize(count = sum(count)) %>%
  left_join(phl_nhood_cent, by=c("nhood_origin"="name")) %>% #join origin nhood to nhood centroid
  left_join(., phl_corr_cent, by=c("corridor_dest"="NAME")) %>% #join destination corridor to corr centroid
  dplyr::rename(., origin.geom = geometry.x,
                        dest.geom = geometry.y) #clean-up columns for clarity

#Convert from tibble to data frame for next step
flows_nhood <- as.data.frame(flows_nhood) 

#split point data into lat and long columns
flows_nhood <- flows_nhood %>% 
  mutate(lat.origin = unlist(map(flows_nhood$origin.geom,1)),
         long.origin = unlist(map(flows_nhood$origin.geom,2)),
         lat.dest = unlist(map(flows_nhood$dest.geom,1)),
         long.dest = unlist(map(flows_nhood$dest.geom,2)),
         id = as.character(c(1:nrow(.))))

#Maps - straight line segment example
flows_nhood %>% 
  filter(corridor_dest == "East Girard" | 
           corridor_dest == 'Navy Yard' | 
           corridor_dest == '5th and Olney' | 
           corridor_dest == 'East Passyunk' |
           corridor_dest == '36th Street and vicinity' |
           corridor_dest == 'Market West') %>% 
  ggplot() + 
  geom_polygon(data = PHL_boundary_unproj, aes(x=X, y=Y), fill = "grey20") + 
  geom_segment(aes(x = lat.origin, y = long.origin, xend = lat.dest, yend = long.dest,  
                   color=count, alpha = count),
               arrow = arrow(length = unit(0.01, "cm")), 
               size = 1,
               lineend = "round") +
  scale_colour_distiller(palette="Reds", name="Count", guide = "colorbar", direction = 1) +
  coord_equal() +
  mapTheme() + 
  facet_wrap(~corridor_dest, ncol = 2) +
  labs(title = "Trips")
```

## Nightlife corridors in Philadelphia

Next, we observe foot traffic along Philadelphia's commercial corridors. This analysis relies on a shapefile from the City of Philadelphia's Planning department which demarcates individual corridors and districts throughout the city. According to the [available metadata](https://metadata.phila.gov/#home/datasetdetails/564236a55737e1f263ae5e3f/representationdetails/56423a4e902dbdd813db9a55/) "locations range from large, regional and specialty destinations to corridors that reflect the evolving economy, culture, and aesthetic traditions of surrounding neighborhoods." This means that the gegoraphies vary in size and character.

Figure X.X below maps the commercial corridors and shows the corridor-wide vacancy rate, which varies from close to 0% to upwards of 60% throughout the city. Smaller corridors in far West Philadelphia, North Philadelphia tend to have the highest vacancy rates.

```{r}
###Corridors
#Load Philadelphia commercial corridors dataset and transform
corridors <- phl_corridors

dat_restaurants <-
dat_hour %>%
  filter(top_category == "Restaurants and Other Eating Places") 

corridors_filter <- corridors %>%
  select(OBJECTID, NAME, GLA, P_DIST, ST_EXT, PT_ADD, VAC_RATE) %>%
  mutate(VAC_RATE = str_remove_all(VAC_RATE, pattern = "%")) %>%
  mutate(VAC_RATE = as.numeric(VAC_RATE)) %>%
  st_as_sf() %>%
  st_transform('ESRI:102271') 
  
# Plot commercial corridors with colors as vacancy rate
ggplot() +
  geom_sf(data = phillyBoundary, fill = "black") +
  geom_sf(data = corridors_filter, aes(fill = VAC_RATE), color = "transparent") +
  scale_fill_viridis_c(option = "plasma") +
  labs(title = "Philadelphia Commercial Corridors") + 
  mapTheme()
```

Figure X.X below breaks down the count of restaunts in a given corridor per square miile.  We find that central corridors and districts tend to have more restaurants per square mile. Other corridors. That said, corridors across the city have a high concentration of restaurants as well.
```{r}
###Restaurants (takes a really long time)
dat_restaurants_grid <- dat_restaurants %>%
  select(geometry) %>%
  na.omit() %>%
  st_as_sf() %>%
  st_transform('ESRI:102271') %>% 
    distinct()

#Merge restaurants with commercial corridors 
corridors_restaurants <- 
  dplyr::select(dat_restaurants_grid) %>% 
  mutate(countRestaurant = 1) %>% 
  aggregate(., corridors_filter, sum) %>%
  mutate(countRestaurant = replace_na(countRestaurant, 0),
         uniqueID = rownames(.),
         area = st_area(geometry) * .00000038610,
         count_per_mile = as.numeric(countRestaurant / area),
         cvID = sample(round(nrow(corridors_filter) / 24), size=nrow(corridors_filter), replace = TRUE))

#Plot map of number of restaurants per square mile
ggplot() +
  geom_sf(data = phillyBoundary, fill = "black") +
  geom_sf(data = corridors_restaurants, aes(fill = count_per_mile), color = "transparent") +
  scale_fill_viridis(trans = "sqrt") +
  labs(title = "Count of Restaurants per Square Mile in Each Commercial Corridor",
       subtitle = "Figure X.X") +
  mapTheme()
```

Below is an animation showing restaurant popularity by corridor.
```{r}
#Create animated map of number of restaurant trips by corridor per hour
dat_restaurants_cord <- dat_restaurants %>%
  st_as_sf() %>%
  st_transform('ESRI:102271') 

dat_corridors_restaurants <-
  st_join(corridors_filter, dat_restaurants_cord, ) %>% 
  group_by(NAME, GLA, Hour) %>%
  summarize(
            Avg_Popularity = mean(Count),
            Total_Visits = sum(Count),
            Med_Dwell_Time = mean(median_dwell)) %>%
  st_as_sf() 

dat_corridors_restaurants <- dat_corridors_restaurants %>%
  mutate(
    GLA = as.numeric(gsub(",", "", GLA)),
    Visits_Per_Area = Total_Visits / GLA * 5)

#Plot of restaurant popularity between 7 and 8pm
# dat_corridors_restaurants %>%
#   subset(Hour == 19) %>% #selecting for nighttime hours
#   ggplot() + 
#   geom_sf(data = phl_cbg, fill = "#440255", color = "transparent") +
#   geom_sf(aes(fill = Visits_Per_Area), color = "transparent") + 
#   scale_fill_viridis(trans = "sqrt") +
#   labs(title = "Total Restaurant Visits by Commercial Corridor") 


#Animation of restaurant popularity by hour
restaurant.corr.animation.data <-
    dat_corridors_restaurants %>%
    st_sf() %>%
    mutate(Pop_String = case_when(Visits_Per_Area < .4 ~ "5",
                              Visits_Per_Area >= .4 & Visits_Per_Area <.8 ~ "4",
                              Visits_Per_Area >= .8 & Visits_Per_Area <1.2 ~ "3",
                              Visits_Per_Area >= 1.2 & Visits_Per_Area <1.6 ~ "2",
                              Visits_Per_Area >= 2 ~ "1")) %>%
    mutate(Pop_String  = fct_relevel(Pop_String, "5","4","3","2","1"))

restaurant_corr_animation <-
  ggplot() +
  geom_sf(data = phl_cbg, fill = "#440255", color = "transparent") +   
  geom_sf(data = restaurant.corr.animation.data, aes(fill = Visits_Per_Area), color = "transparent") +
     scale_fill_viridis(trans = "sqrt") +
    labs(title = "Restaurant Popularity by Hour",
         subtitle = "One Hour Intervals: {current_frame}") +
  theme(panel.background = element_rect(fill = "black"),
         panel.grid.major = element_line(color = "transparent"),
          panel.grid.minor = element_line(colour = "transparent")) +
    transition_manual(Hour)

animate(restaurant_corr_animation, duration=20, renderer = gifski_renderer())
```

Figure X.X shows the number of bars in each commercial corridor per square mile. Again, we see the corridors around Center City generally showing a higher concentration of bars.
```{r Bars by corridor}
###Bars
dat_bars_grid <- dat_bars %>%
  select(geometry) %>%
  #na.omit() %>%
  st_as_sf() %>%
  st_transform('ESRI:102271') %>%
    distinct()

###Bars
corridors_bars <- 
  dplyr::select(dat_bars_grid) %>% 
  mutate(countBar = 1) %>% 
  aggregate(., corridors_filter, sum) %>%
  mutate(countBar = replace_na(countBar, 0),
         uniqueID = rownames(.),
         area = st_area(geometry) * .00000038610,
         count_per_mile = as.numeric(countBar / area),
         cvID = sample(round(nrow(corridors_filter) / 24), size=nrow(corridors_filter), replace = TRUE))

#Plot map of number of bars per square mile
ggplot() +
  geom_sf(data = phillyBoundary, fill = "black") +
  geom_sf(data = corridors_bars, aes(fill = count_per_mile), color = "transparent") +
  scale_fill_viridis(trans = "sqrt") +
  labs(title = "Count of Bars per Square Mile in Each Commercial Corridor", 
       subtitle = "Figure X.X") +
  mapTheme()
```

Below is an animation showing bar popularity aggregated by commercial corridor.
```{r bars animation}
#Create animated map of number of bar trips by corridor per hour
dat_bars_cord <- dat_bars %>%
  st_as_sf() %>%
  st_transform('ESRI:102271') 

dat_corridors_bars <-
  st_join(corridors_filter, dat_bars_cord, ) %>% 
  group_by(NAME, GLA, Hour) %>%
  summarize(Avg_Popularity = mean(Count),
            Total_Visits = sum(Count),
            Med_Dwell_Time = mean(median_dwell)) %>%
  st_as_sf() 

dat_corridors_bars <- dat_corridors_bars %>%
  mutate(
    GLA = as.numeric(gsub(",", "", GLA)),
    Visits_Per_Area = Total_Visits / GLA * 5)

#Plot of bar popularity between 7 and 8pm
# dat_corridors_bars %>%
#   subset(Hour == 19) %>% #selecting for nighttime hours
#   ggplot() + 
#   geom_sf(data = phl_cbg, fill = "#440255", color = "transparent") +
#   geom_sf(aes(fill = Visits_Per_Area), color = "transparent") + 
#   scale_fill_viridis(trans = "sqrt") +
#   labs(title = "Total Bar Visits by Commercial Corridor") 

#Animation of bar popularity by hour
bar.corr.animation.data <-
    dat_corridors_bars %>%
    st_sf() %>%
    mutate(Pop_String = case_when(Visits_Per_Area < .4 ~ "5",
                              Visits_Per_Area >= .4 & Visits_Per_Area <.8 ~ "4",
                              Visits_Per_Area >= .8 & Visits_Per_Area <1.2 ~ "3",
                              Visits_Per_Area >= 1.2 & Visits_Per_Area <1.6 ~ "2",
                              Visits_Per_Area >= 2 ~ "1")) %>%
    mutate(Pop_String  = fct_relevel(Pop_String, "5","4","3","2","1"))

bar_corr_animation <-
  ggplot() +
  geom_sf(data = phl_cbg, fill = "#440255", color = "transparent") +   
  geom_sf(data = bar.corr.animation.data, aes(fill = Visits_Per_Area), color = "transparent") +
     scale_fill_viridis(trans = "sqrt") +
    labs(title = "Bar Popularity by Hour",
         subtitle = "One Hour Intervals: {current_frame}") +
  theme(panel.background = element_rect(fill = "black"),
         panel.grid.major = element_line(color = "transparent"),
          panel.grid.minor = element_line(colour = "transparent")) +
    transition_manual(Hour)

animate(bar_corr_animation, duration=20, renderer = gifski_renderer())
```

## Visit duration to bars and restaurants in Philadelphia

The following Figures look at the median dwell time by bars and restaurants per commercial corridor.
```{r}
#Restaurant data
dat_corridors_restaurants <-
  st_join(corridors_filter, dat_restaurants_cord, ) %>% 
  group_by(NAME, GLA, Hour) %>%
  summarize(
            Avg_Popularity = mean(Count),
            Total_Visits = sum(Count),
            Med_Dwell_Time = mean(median_dwell)) %>%
  st_as_sf() 

dat_corridors_restaurants <- dat_corridors_restaurants %>%
  mutate(
    GLA = as.numeric(gsub(",", "", GLA)),
    Visits_Per_Area = Total_Visits / GLA * 5)

#Plot of restaurant median dwell time 
dat_corridors_restaurants %>%
  subset(Hour == 19) %>% #selecting for nighttime hours
  ggplot() + 
  geom_sf(data = phl_cbg, fill = "light gray", color = "white") +
  geom_sf(aes(fill = Med_Dwell_Time), color = "transparent") + 
  scale_fill_viridis(trans = "sqrt") +
  labs(title = "Median Restaurant Dwell Time by Commercial Corridor",
       subtitle = "Figure X.X") +
  mapTheme()

#Plot of bar median dwell time 
dat_corridors_bars %>%
  subset(Hour == 19) %>% #selecting for nighttime hours
  ggplot() + 
  geom_sf(data = phl_cbg, fill = "light gray", color = "white") +
  geom_sf(aes(fill = Med_Dwell_Time), color = "transparent") + 
  scale_fill_viridis(trans = "sqrt") +
  labs(title = "Median Bar Dwell Time by Commercial Corridor",
       subtitle = "Figure X.X") +
  mapTheme()
```

## Neighborhood and corridor comparisons

Finally, we began exploring businesses at the corridor level and understand how different corridors attract different visitors from across the city. We plan to systematically expand the scope of this study to other corridors across the city. We also believe that this data is particularly relevant to business and economic development professionals in the COVID-19 recovery efforts.

Using the Commercial Corridor shapefile available on Open Data Philly, we selected two corridors to compare:  East Girard Avenue, a central corridor in the Fishtown neighborhood, and West Girard Avenue in Brewerytown. 

```{r East Girard data wrangling}
#East Girard Corridor
EastGirard_corr <- 
  phl_corridors %>% 
  filter(NAME == "East Girard") %>% 
  st_as_sf()

dat_EGC <- dat2[EastGirard_corr,]

dat_EGC_cbg <- 
  dat_EGC %>%
  select(safegraph_place_id, 
         date_range_start, 
         top_category, 
         sub_category, 
         visitor_home_cbgs, 
         geometry) %>%
  mutate(visitor_home_cbgs = str_remove_all(visitor_home_cbgs, pattern = "\\[|\\]")) %>%
  mutate(visitor_home_cbgs = str_remove_all(visitor_home_cbgs, pattern = "\\{|\\}")) %>%
  mutate(visitor_home_cbgs = str_remove_all(visitor_home_cbgs, pattern = '\\"|\\"')) %>%
  mutate(visitor_home_cbgs = str_split(visitor_home_cbgs, pattern = ",")) %>%
  unnest(visitor_home_cbgs) %>%
  separate(.,
           visitor_home_cbgs,
           c("cbg_origin", "visitors"),
           sep = ":") %>%
  mutate(cbg_origin = as.numeric(cbg_origin),
         visitors = as.numeric(visitors)) %>%
  filter(top_category == "Drinking Places (Alcoholic Beverages)" |
           top_category == "Restaurants and Other Eating Places" |
           top_category == "Traveler Accommodation" |
           top_category == "Gambling Industries" |
           top_category == "Promoters of Performing Arts, Sports, and Similar Events" |
           top_category == "Performing Arts Companies") %>%
    mutate(corridor = "East Girard")

#West Girard
WestGirard_corr <- 
  phl_corridors %>% 
  filter(NAME == "West Girard") %>% 
  st_as_sf()

dat_WGC <- dat2[WestGirard_corr,]

dat_WGC_cbg <- 
  dat_WGC %>%
  select(safegraph_place_id, 
         date_range_start, 
         top_category, 
         sub_category, 
         visitor_home_cbgs, 
         geometry) %>%
  mutate(visitor_home_cbgs = str_remove_all(visitor_home_cbgs, pattern = "\\[|\\]")) %>%
  mutate(visitor_home_cbgs = str_remove_all(visitor_home_cbgs, pattern = "\\{|\\}")) %>%
  mutate(visitor_home_cbgs = str_remove_all(visitor_home_cbgs, pattern = '\\"|\\"')) %>%
  mutate(visitor_home_cbgs = str_split(visitor_home_cbgs, pattern = ",")) %>%
  unnest(visitor_home_cbgs) %>%
  separate(.,
           visitor_home_cbgs,
           c("cbg_origin", "visitors"),
           sep = ":") %>%
  mutate(cbg_origin = as.numeric(cbg_origin),
         visitors = as.numeric(visitors)) %>%
  filter(top_category == "Drinking Places (Alcoholic Beverages)" |
           top_category == "Restaurants and Other Eating Places" |
           top_category == "Traveler Accommodation" |
           top_category == "Gambling Industries" |
           top_category == "Promoters of Performing Arts, Sports, and Similar Events" |
           top_category == "Performing Arts Companies") %>%
  mutate(corridor = "West Girard")
```

Figures X.Xa and X.Xb show the volume of visitors to the East Girard Corridor (outlined in red) by census block group. While visitors from across the city frequent East Girard corridor, particularly for its restaurants, Figure X.Xb clarifies that the corridor mostly caters to a local customer base of surrounding census tracts.

Block groups shaded gray indicate that no resident made a trip to this corridor.
```{r}
dat_EGC_cbg %>%
  dplyr::group_by(cbg_origin, top_category) %>%
  dplyr::summarize(visitors = sum(visitors)) %>%
  st_drop_geometry() %>%
  dplyr::rename(., GEOID10 = cbg_origin) %>%
  left_join(phl_cbg, by = "GEOID10") %>%
  st_as_sf() %>%
  drop_na(geometry) %>% 
  ggplot() +
  geom_sf(data = phl_cbg, fill = "grey70", color = "transparent") +
  geom_sf(aes(fill = q5(visitors), geometry = geometry), color = "transparent") + 
  geom_sf(data = phl_nhoods, color = "white", fill = "transparent") +
  geom_sf(data = EastGirard_corr, color = "red", fill = "transparent", lwd = 1) +
  scale_fill_manual(values = palette5,
                    aesthetics = c("colour", "fill"),
                    name = "Visitor Count \n(Quintile)") +
  mapTheme() +
  labs(title = "East Girard Restaurants: Where do visitors come from?",
       subtitle = "Figure X.Xa") +
  facet_wrap(~top_category)

dat_EGC_cbg %>%
  dplyr::group_by(cbg_origin, top_category) %>%
  dplyr::summarize(visitors = sum(visitors)) %>%
  st_drop_geometry() %>%
  dplyr::rename(., GEOID10 = cbg_origin) %>%
  left_join(phl_cbg, by = "GEOID10") %>%
  st_as_sf() %>%
  drop_na(geometry) %>% 
  ggplot() +
  geom_sf(data = phl_cbg, fill = "grey70", color = "transparent") +
  geom_sf(aes(fill = visitors, geometry = geometry), color = "transparent") + 
  geom_sf(data = phl_nhoods, color = "white", fill = "transparent") +
  geom_sf(data = EastGirard_corr, color = "red", fill = "transparent", lwd = .5) +
  scale_fill_viridis() +
  mapTheme() +
  labs(title = "East Girard Restaurants: Where do visitors come from?",
       subtitle = "Figure X.Xb") +
  facet_wrap(~top_category)
```

Figures 3.12a and 3.12b show the volume of visitors to West Girard corridor. While the corridor appears to draw visitors from fewer census tracts overall, the corridor draws from a similarly local customer base of the surrounding block groups.

As before, block groups shaded gray have no residents that made trips to this corridor.
```{r West Girard viz}
dat_WGC_cbg %>%
  dplyr::group_by(cbg_origin, top_category) %>%
  dplyr::summarize(visitors = sum(visitors)) %>%
  st_drop_geometry() %>%
  dplyr::rename(., GEOID10 = cbg_origin) %>%
  left_join(phl_cbg, by = "GEOID10") %>%
  st_as_sf() %>%
  drop_na(geometry) %>% 
  ggplot() +
  geom_sf(data = phl_cbg, fill = "grey70", color = "transparent") +
  geom_sf(aes(fill = q5(visitors), geometry = geometry), color = "transparent") + 
  geom_sf(data = phl_nhoods, color = "white", fill = "transparent") +
  geom_sf(data = WestGirard_corr, color = "red", fill = "transparent", lwd = 1) +
  scale_fill_manual(values = palette5,
                    aesthetics = c("colour", "fill"),
                    name = "Visitor Count \n(Quintile)") +
  mapTheme() +
  labs(title = "West Girard: Where do visitors come from?",
       subtitle = "Figure X.Xa") +
  facet_wrap(~top_category)

dat_WGC_cbg %>%
  dplyr::group_by(cbg_origin, top_category) %>%
  dplyr::summarize(visitors = sum(visitors)) %>%
  st_drop_geometry() %>%
  dplyr::rename(., GEOID10 = cbg_origin) %>%
  left_join(phl_cbg, by = "GEOID10") %>%
  st_as_sf() %>%
  drop_na(geometry) %>% 
  ggplot() +
  geom_sf(data = phl_cbg, fill = "grey70", color = "transparent") +
  geom_sf(aes(fill = visitors, geometry = geometry), color = "transparent") + 
  geom_sf(data = phl_nhoods, color = "white", fill = "transparent") +
  geom_sf(data = WestGirard_corr, color = "red", fill = "transparent", lwd = .5) +
  scale_fill_viridis() +
  mapTheme() +
  labs(title = "East Girard Restaurants: Where do visitors come from?",
       subtitle = "Figure X.Xb") +
  facet_wrap(~top_category)
```

We can further explore the individual corridor analysis by tying the trip origins to census data. This allows us to construct a customer profile of the people visiting specific corridors throughout the city.
```{r census pull, options(tigris_use_cache = TRUE)}
phl_blockgroups <- 
  get_acs(geography = "block group", 
        variables = c("B01003_001E", 
                      "B02001_002E", 
                      "B01002_001E",
                      "B19013_001E", 
                      "B25064_001E"),
        year=2018, 
        state=42, 
        county=101, 
        geometry=T, 
        output = "wide") %>%
  st_transform('ESRI:102728')%>%
  rename(TotalPop = B01003_001E,
         Whites = B02001_002E,
         MedAge = B01002_001E,
         MedHHInc = B19013_001E,
         MedRent = B25064_001E,
         GEOID10 = GEOID) %>%
  dplyr::select(-NAME, -starts_with("B")) %>%
  mutate(pctWhite = ifelse(TotalPop > 0, Whites / TotalPop,0),
         GEOID10 = as.numeric(GEOID10)) %>%
  dplyr::select(-Whites) 
```

The below Figure X.X compares demographic indicators pulled from the census across the two customer bases. East Girard corridor's customer base, shown in purple, tend to come from whiter census tracts with a higher median household income and median rent. West Girard corridor's customer base tend to come from less white census tracts with lower incomes and median rent. Both corridors, however, pull from census tracts with a similar median age.
```{r corridor comparison}
rbind(dat_EGC_cbg, dat_WGC_cbg) %>%
  rename(GEOID10 = cbg_origin) %>%
  dplyr::select(GEOID10, visitors, corridor) %>%
  st_drop_geometry() %>%
  left_join(phl_blockgroups, by = "GEOID10") %>%
  pivot_longer(.,
               cols = c("MedAge", "MedHHInc", "MedRent", "pctWhite"),
               names_to = "variable") %>%
  dplyr::select(GEOID10, corridor, visitors, variable, value) %>%
  ggplot(., aes(x = value, fill = corridor, weight = visitors)) + 
  geom_density(alpha = .5) +
  scale_fill_manual(values = palette3,
                    aesthetics = c("colour", "fill"),
                    name = "Visitor Count \n(Quintile)") +
  facet_wrap( ~ variable, scales = "free") +
  plotTheme() +
  labs(title = 'East Girard vs. West Girard Corridors',
       subtitle = "Figure X.X")
```

# 4. Prediction

Our model will predict the number of trips to commercial corridors given features of the corridor such as retail mix, amenities, location, etc. Ultimately, we hope this will be a scenario-based tool that allows users to adjust the business mix for a given corriodr as a way to understand how different economic development strategies will impact popularity.

While the model will predict the number of trips, the ultimate output for the user will be relative popularity of the corridor compared to the average. This is becasue the data is inherently noisy and we cannot be confident that a raw trip count is accurate.

This first code block loads in data that we will use for the predictive model.
```{r loading in prediction data}

library(FNN)
nn_function <- function(measureFrom,measureTo,k) {
  measureFrom_Matrix <- as.matrix(measureFrom)
  measureTo_Matrix <- as.matrix(measureTo)
  nn <-   
    get.knnx(measureTo, measureFrom, k)$nn.dist
  output <-
    as.data.frame(nn) %>%
    rownames_to_column(var = "thisPoint") %>%
    gather(points, point_distance, V1:ncol(.)) %>%
    arrange(as.numeric(thisPoint)) %>%
    group_by(thisPoint) %>%
    dplyr::summarize(pointDistance = mean(point_distance)) %>%
    arrange(as.numeric(thisPoint)) %>% 
    dplyr::select(-thisPoint) %>%
    pull()
  
  return(output)  
}


#Corridor shapefile
phl_corridors_pred <- st_read("http://data.phl.opendata.arcgis.com/datasets/f43e5f92d34e41249e7a11f269792d11_0.geojson", quiet = TRUE)%>%
  st_transform('ESRI:102728') %>%
  select(4:5,18, 'Shape__Area') %>%
  rename(., corridor = NAME,
         district = P_DIST,
         vacancy = VAC_RATE,
         area = Shape__Area) %>%
  mutate()

#Spatial Features from SafeGraph Data
bar.sf <- dat2 %>%
  filter(top_category == "Drinking Places (Alcoholic Beverages)") %>%
  select(location_name) %>%
  st_as_sf(coords = "geometry", crs = 4326, agr = "constant")

restaurant.sf <- dat2 %>%
  filter(top_category == "Restaurants and Other Eating Places") %>%
  select(location_name) %>%
  st_as_sf()

arts.sf <- dat2 %>%
  filter(top_category == "Promoters of Performing Arts, Sports, and Similar Events" |
           top_category == "Performing Arts Companies") %>%
  select(location_name) %>%
  st_as_sf()

college.sf <- dat2 %>%
  filter(top_category == "Colleges, Universities, and Professional Schools") %>%
  select(location_name) %>%
  st_as_sf()

sports.sf <- dat2 %>%
  filter(top_category == "Spectator Sports") %>%
  select(location_name) %>%
  st_as_sf()

casinos.sf <- dat2 %>%
  filter(sub_category == "Casino Hotels") %>%
  select(location_name) %>%
  st_as_sf()

hotels.sf <- dat2 %>%
  filter(sub_category == "Hotels (except Casino Hotels) and Motels") %>%
  select(location_name) %>%
  st_as_sf()

st_c <- st_coordinates

dat_corr <-
  dat2 %>% 
  st_join(phl_corridors_pred, join = st_intersects) %>%
  # st_join(phl_nhoods_pred, join = st_intersects) %>%
  drop_na(corridor)
```

Next, we generate features for our model. These features include counts of business types and tags (such as "late night"), nearest neighbor variables, and distance to specific locations.


```{r feature engineering for distance to transit}

septaStops <- 
  rbind(
    st_read("https://opendata.arcgis.com/datasets/8c6e2575c8ad46eb887e6bb35825e1a6_0.geojson") %>% 
    mutate(Line = "El") %>%
    select(Station, Line),
      st_read("https://opendata.arcgis.com/datasets/2e9037fd5bef406488ffe5bb67d21312_0.geojson") %>%
      mutate(Line ="Broad_St") %>%
      select(Station, Line)) %>%
   st_transform(st_crs(phl_blockgroups))
```


```{r feature engineering}
dat_corr <-
  dat_corr %>%
  mutate(total = 1,
         bars = ifelse(top_category == 'Drinking Places (Alcoholic Beverages)', 1, 0),
         restaurant = ifelse(top_category == 'Restaurants and Other Eating Places', 1, 0),
         arts = ifelse(top_category == 'Promoters of Performing Arts, Sports, and Similar Events' |
                         top_category == 'Performing Arts Companies', 1, 0),
         restaurant = ifelse(top_category == 'Restaurants and Other Eating Places', 1, 0),
         jrcol = ifelse(top_category == 'Junior Colleges', 1, 0),
         college = ifelse(top_category == 'Colleges, Universities, and Professional Schools', 1, 0),
         sports = ifelse(top_category == 'Spectator Sports', 1, 0),
         museum = ifelse(top_category == 'Museums, Historical Sites, and Similar Institutions', 1, 0),
         hotels = ifelse(top_category == 'Traveler Accommodation', 1, 0),
         bars_nn = nn_function(st_c(st_centroid(dat_corr)), st_c(st_centroid(bar.sf)), 4),
         rest_nn = nn_function(st_c(st_centroid(dat_corr)), st_c(st_centroid(restaurant.sf)), 4),
         arts_nn = nn_function(st_c(st_centroid(dat_corr)), st_c(st_centroid(arts.sf)), 4),
         college_nn = nn_function(st_c(st_centroid(dat_corr)), st_c(st_centroid(college.sf)), 1),
         sports_nn = nn_function(st_c(st_centroid(dat_corr)), st_c(st_centroid(sports.sf)), 1),
         hotels_nn = nn_function(st_c(st_centroid(dat_corr)), st_c(st_centroid(hotels.sf)), 4),
         casinos_nn = nn_function(st_c(st_centroid(dat_corr)), st_c(st_centroid(casinos.sf)), 1),
         transit_nn = nn_function(st_c(st_centroid(dat_corr)), st_c(st_centroid(septaStops)), 1),
         late_night = if_else(grepl("Late Night", dat_corr$category_tags), 1, 0),
         bar_pub = if_else(grepl("Bar or Pub", dat_corr$category_tags), 1, 0))
```


```{r census feature engineering}
phl_blockgroups <- 
  get_acs(geography = "block group", 
        variables = c("B01003_001E", 
                      "B02001_002E", 
                      "B01002_001E",
                      "B19013_001E", 
                      "B25064_001E"),
        year=2018, 
        state=42, 
        county=101, 
        geometry=T, 
        output = "wide") %>%
  st_transform('ESRI:102728')%>%
  rename(TotalPop = B01003_001E,
         Whites = B02001_002E,
         MedAge = B01002_001E,
         MedHHInc = B19013_001E,
         MedRent = B25064_001E,
         GEOID10 = GEOID) %>%
  mutate(pctWhites = ((Whites / TotalPop)*100),
         raceContext = ifelse(pctWhites > .5, "Majority White", "Majority Non-White"),
         incomeContext = ifelse(MedHHInc > 46116, "High Income", "Low Income"),
         ageContext = ifelse(MedAge > 34.5, "High Median Age", "Low Median Age"),
         rentContext = ifelse(MedRent > 1032, "High Median Rent", "Low Median Rent"))

```

```{r import weather data for feature engineering}

# Get 2018 weather data
weather.Panel <- 
  riem_measures(station = "ORD", date_start = "2018-01-01", date_end = "2018-12-31") %>%
  dplyr::select(valid, tmpf, p01i, sknt)%>%
  replace(is.na(.), 0) %>%
    mutate(date_range_start = as.Date(ymd_h(substr(valid,1,13)))) %>%
    mutate(week = week(date_range_start),
           dotw = wday(date_range_start, label=TRUE)) %>%
    group_by(date_range_start) %>%
    summarize(Temperature = max(tmpf),
              Precipitation = sum(p01i),
              Wind_Speed = max(sknt)) %>%
    mutate(Temperature = ifelse(Temperature == 0, 42, Temperature))
```



Next we wrangle the data to aggregate it at the corridor level. To start, we are looking at aggregated SafeGraph variables, such as median dwell time and average distance from home. 

We also review the share of certain business types and characteristics along a specific corridor. These variables count the SafeGraph points of interest that fit into given category (bars, restaurants, college) and divide it by hte total number of businesses along the corridor. We also divide by 12 since the data is disaggregated by month.

Finally, this code generates nearest neighbor variables for business by corridor.

```{r Observe correlations}
corr_test <- 
  dat_corr %>%
  group_by(corridor, area) %>%
  summarize(visits = sum(raw_visit_counts),
            total = sum(total),
            area = mean(area),
            sg_visitors = sum(raw_visitor_counts),
            sg_dwell = mean(median_dwell, na.rm = TRUE),
            sg_distance_home = mean(distance_from_home, na.rm = TRUE),
            share_bars = sum(bars, na.rm = TRUE)/12/area,
            share_rest = sum(restaurant, na.rm = TRUE)/12/area,
            share_arts = sum(arts, na.rm = TRUE)/12/area,
            share_jrcol = sum(jrcol, na.rm = TRUE)/12/area,
            share_college = sum(college, na.rm = TRUE)/12/area,
            share_sports = sum(sports, na.rm = TRUE)/12/area,
            share_museums = sum(museum, na.rm = TRUE)/12/area,
            nn_bars = mean(bars_nn),
            nn_rest = mean(rest_nn),
            nn_arts = mean(arts_nn),
            nn_college = mean(college_nn),
            nn_sports = mean(sports_nn),
            nn_casinos = mean(casinos_nn),
            nn_hotels = mean(hotels_nn),
            share_late_tag = sum(late_night, na.rm = TRUE)/12/area,
            share_barpub_tag = sum(bar_pub, na.rm = TRUE)/12/area)
```

The following correlation plots summarize our findings. Looking first at the SafeGraph summary variables, the distnace from home has a strong relationship with total visits. The count of visitors also demonstrates a strong relationship, but this is likely collinear with the the number of visits.
```{r corr plots}
#Safegraph summary variables
corr_test %>% 
  st_drop_geometry() %>%
  pivot_longer(., cols = starts_with("sg_"), names_to = "variable", values_to="value") %>%
  ggplot(aes(value, visits)) +
  geom_point(size = .5) + 
  geom_smooth(method = "lm", se=F, colour = "#FA7800") +
  facet_wrap(~variable, ncol = 3, scales = "free") +
  labs(title = "Visits as a function of SafeGraph SUmmary Variables",
       subtitle = "Figure X.X") +
  plotTheme()
```

The next set of plots show the relationship between the share a specific business type by corridor and the number of trips. There seems to be a relationship with businesses that could be classified as anchor institutions, such as colleges and museums, and the number of trips. These corrplots also look at businesses with a specific tag, such as a "late night" or "bar and pub". These corrplots rely on the businesses recorded in the SafeGraph dataset and represent a sample. Do we want to use another, more comprehensive dataset instead (e.g. OSM)?

```{r}
#Count variables
corr_test %>% 
  st_drop_geometry() %>%
  pivot_longer(., cols = starts_with("share_"), names_to = "variable", values_to="value") %>%
  ggplot(aes(value, visits)) +
  geom_point(size = .5) + 
  geom_smooth(method = "lm", se=F, colour = "#FA7800") +
  facet_wrap(~variable, ncol = 3, scales = "free") +
  labs(title = "Visits as a function of Continuous Count Variables",
       subtitle = "Figure X.X") +
  plotTheme()
```

Finally, we've generated nearest neighbor variables for the same businesses. So far, these are not demonstrating a particualrly strong relationship with foot traffic.
```{r}
#Nearest neighbor variables
corr_test %>% 
  st_drop_geometry() %>%
  pivot_longer(., cols = starts_with("nn_"), names_to = "variable", values_to="value") %>%
  ggplot(aes(value, visits)) +
  geom_point(size = .5) + 
  geom_smooth(method = "lm", se=F, colour = "#FA7800") +
  facet_wrap(~variable, ncol = 3, scales = "free") +
  labs(title = "Visits as a function of Nearest Neighbor Variables",
       subtitle = "Figure X.X") +
  plotTheme()
```

# 5. Next Steps

For next week, we will primarily focus on preparing our presentation by doing the following:

* Distilling our EDA presentaiton from last time and incorporate COVID-19 data.
* Present features that correlate with corridor trips.
* Refine wire frame.