---
title: "MUSA Practicum Nighttime Economy: Exploratory Analysis"
author: "Maddy Kornhauser, Brian Rawn, Sabrina Lee"
date: "2/22/2021"
output:
  github_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	out.height = "\\textheight",  out.width = "\\textwidth"
)
```
# 1. Use Case Development

How can we help business improvement districts (BIDs) and small businesses better understand who is visiting their commercial corridor for nightlife purposes to further economic recovery efforts following the COVID-19 pandemic? 

* *Approach*: Use safegraph data to understand how nightlife patterns differ across commercial corridors in the city. This analysis would allow BIDs and business to understand who is coming to their area for nightlife (avg. income, demographics, and from which areas), where they are coming from, where and when they are visiting, and how these patterns changed before and after COVID-19. 

* *Potential Deliverable*: An interactive economic development tool to help BIDs, small businesses, and other public stakeholders understand how they size up against other commercial corridors in the city and gain basic statistic on their customer base.

# 2. The SafeGraph Dataset

Safegraph uses anonymized cell phone GPS data to record trips to commercial points of interest. This data can tell us from where a trip was made, what time the trip was made, and how long the individual stated at the point of interest.

* Pros: 
  + Brand new dataset. 
  + Lots of unexplored applications and potential insights

* Cons: 
  + Data is new and has a significant amount of incorrectly attributed trips, especially in urban areas

# 3. Exploratory Data Analysis

```{r load packages, include=FALSE}
library(tidyverse)
library(tidycensus)
library(sf)
library(lubridate)
library(datetime)
library(viridis)
library(gridExtra)
library(kableExtra)
library(dplyr)
library(ggplot2)
library(gganimate)


qBr <- function(df, variable, rnd) {
  if (missing(rnd)) {
    as.character(quantile(round(df[[variable]],0),
                          c(.01,.2,.4,.6,.8), na.rm=T))
  } else if (rnd == FALSE | rnd == F) {
    as.character(formatC(quantile(df[[variable]]), digits = 3),
                 c(.01,.2,.4,.6,.8), na.rm=T)
  }
}
q5 <- function(variable) {as.factor(ntile(variable, 5))}
palette3 <- viridis_pal()(3)
palette5 <- viridis_pal()(5)
mapTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 14,colour = "black"),
    plot.subtitle=element_text(face="italic"),
    plot.caption=element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),axis.title = element_blank(),
    axis.text = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=2)
  )
}

plotTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 16,colour = "black"),
    plot.subtitle = element_text(face="italic"),
    plot.caption = element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    panel.grid.major = element_line("grey80", size = 0.1),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=2),
    strip.background = element_rect(fill = "grey80", color = "white"),
    strip.text = element_text(size=12),
    axis.title = element_text(size=12),
    axis.text = element_text(size=10),
    plot.background = element_blank(),
    legend.background = element_blank(),
    legend.title = element_text(colour = "black", face = "italic"),
    legend.text = element_text(colour = "black", face = "italic"),
    strip.text.x = element_text(size = 14)
  )
}


setwd("~/GitHub/musa_practicum_nighttime")
```

```{r load & wrangle data}
dat <- read.csv("./data/moves_2018.csv")
phila <- st_read("./demo/phila.geojson") %>%
  st_transform('ESRI:102728') %>%
  st_as_sf()

dat2 <- dat %>% 
  dplyr::select(safegraph_place_id, 
                date_range_start, 
                date_range_end, 
                raw_visit_counts,
                raw_visitor_counts, 
                visits_by_day, 
                poi_cbg, 
                visitor_home_cbgs, 
                visitor_daytime_cbgs, 
                visitor_work_cbgs, 
                visitor_country_of_origin,
                distance_from_home, 
                median_dwell, 
                bucketed_dwell_times, 
                related_same_day_brand, 
                related_same_month_brand, 
                popularity_by_hour, 
                popularity_by_day, 
                device_type) %>%
  left_join(., phila, by = "safegraph_place_id") %>% 
  st_as_sf()

phl_cbg <- st_read("http://data.phl.opendata.arcgis.com/datasets/2f982bada233478ea0100528227febce_0.geojson") %>%
  st_transform('ESRI:102728') %>%
  mutate(GEOID10 = as.numeric(GEOID10))
phl_zip <- st_read("http://data.phl.opendata.arcgis.com/datasets/b54ec5210cee41c3a884c9086f7af1be_0.geojson") %>%
  st_transform('ESRI:102728') %>%
  mutate(CODE = as.numeric(CODE))
phl_nhoods <- 
  st_read("https://raw.githubusercontent.com/azavea/geo-data/master/Neighborhoods_Philadelphia/Neighborhoods_Philadelphia.geojson") %>%
  st_transform('ESRI:102728') %>% 
  st_as_sf()
phl_corridors <- st_read("http://data.phl.opendata.arcgis.com/datasets/f43e5f92d34e41249e7a11f269792d11_0.geojson") %>%
  st_transform('ESRI:102728')
```

## Philadelphia nightlife establishments

Our first research question is where Philadelphia nightlife establishments are located across the city. The following maps indicate where businesses that contribute to the city's nightlife economy are located. The categories include:

* Bars (Drinking Places)
* Restaurants 
* Hotels (Travel Accomodation)
* Casinos (Gambling Industries)
* Arts Venues (Promotes of Performing Arts)

Figure 3.1 below shows the spatial patterns of the business categories. Bars and restaurants represent the highest number of businesses which are spread across the city. Hotels are mostly clustered in the central district of the city and near the airport in the southwest portion of the city. There are far fewer casinos and arts venues.

Throughout the analysis, we pay special attention to bars and restaurants, as these organizations are well distributed throughout the city and apply to a local Philadelphia customer base.

```{r establishment locations}
dat2 %>%
  filter(top_category == "Drinking Places (Alcoholic Beverages)" |
           top_category == "Restaurants and Other Eating Places" |
           top_category == "Traveler Accommodation" |
           top_category == "Gambling Industries" |
           top_category == "Promoters of Performing Arts, Sports, and Similar Events" |
           top_category == "Performing Arts Companies") %>%
  ggplot() + 
  geom_sf(data = phl_cbg, fill = "grey80", color = "transparent") +
  geom_sf(color = "red", size = .1) +
  labs(title = "Location of Nightlife Establishments",
       subtitle = "Figure 3.1") +
  facet_wrap(~top_category) +
  mapTheme()
```

To look at the spatial patterns another way, we review the distribution of businesses  with a fishnet grid. The fishnet allows us to visualize clusters and hotspots. Starting first with restaurants, Figure 3.2 below demonstrates the spatial patterns of thsee businesses.  Though restaurants are well distributed throughout the city, there appears to be a higher concentration of yellow cells (indicating a higher count of restaurants) in the central part of the city. This corresponds to what we observe in the point data analysis. 
```{r unnesting hour dataset}
#Unnesting popularity by hour variable
dat_hour <- 
  dat2 %>% 
  select(safegraph_place_id, 
         top_category, 
         sub_category, 
         popularity_by_hour, 
         poi_cbg) %>% #select variables
  mutate(popularity_by_hour = str_remove_all(popularity_by_hour, pattern = "\\[|\\]")) %>%   unnest(popularity_by_hour) %>% #unnest values
  separate(.,
           popularity_by_hour,
           c("0", "1", "2", "3", "4", "5", "6", 
             "7", "8", "9", "10", "11", "12", 
             "13", "14", "15", "16", "17", "18",
             "19", "20", "21", "22", "23"),
           sep = ",") %>% #separate into columns by hour
  pivot_longer(cols = 4:27,
               names_to = "Hour",
               values_to = "Visitors") %>% #pivot into long format
  mutate(Hour = as.numeric(Hour),
         Visitors = as.numeric(Visitors))

#Creating a fishnet
phillyBoundary <- 
  phl_zip %>%
  select(geometry) %>%
  st_union() %>%
   st_transform('ESRI:102728') %>% 
  st_as_sf() 

fishnet <- 
  st_make_grid(phillyBoundary, cellsize = 1500, square = FALSE) %>%
  .[phillyBoundary] %>% #MK added this line
  st_sf() %>%
  mutate(uniqueID = rownames(.))

# #Plot fishnet
# ggplot() +
#   geom_sf(data = fishnet, fill = "#440255", color = "black")

#Filter restaurants
restaurants <- dat2 %>%
  filter(top_category == "Restaurants and Other Eating Places")%>%
    st_transform('ESRI:102728')%>%
  mutate(Legend = "Restaurants")

#aggregate restaurant count by fishnet cell
restaurants_net <-
  dplyr::select(restaurants) %>% 
  mutate(countRestaurants = 1) %>% 
  aggregate(., fishnet, sum) %>%
  mutate(countRestaurants = replace_na(countRestaurants, 0),
         uniqueID = rownames(.),
         cvID = sample(round(nrow(fishnet) / 24), 
                       size=nrow(fishnet), replace = TRUE))

#Plot restaurant fishnet
ggplot() +
  geom_sf(data = restaurants_net, aes(fill = countRestaurants), color = NA) +
  scale_fill_viridis() +
  labs(title = "Count of Restaurants for Fishnet", 
       subtitle = "Figure 3.2") +
  mapTheme()
```

Similar to restaurants, Figure 3.3 show that bars are also highly concentrated around Center City.
```{r bars grid}
bars <- dat2 %>%
  filter(top_category == "Drinking Places (Alcoholic Beverages)")%>%
    st_transform('ESRI:102728')%>%
  mutate(Legend = "Bars")

bars_net <-
  dplyr::select(bars) %>% 
  mutate(countBars = 1) %>% 
  aggregate(., fishnet, sum) %>%
  mutate(countBars = replace_na(countBars, 0),
         uniqueID = rownames(.),
         cvID = sample(round(nrow(fishnet) / 24), 
                       size=nrow(fishnet), replace = TRUE)) %>%
  mutate(Legend = "Bars")

ggplot() +
  geom_sf(data = bars_net, aes(fill = countBars), color = NA) +
  scale_fill_viridis() +
  labs(title = "Count of Bars per Fishnet",
       subtitle = "Figure 3.3") +
  mapTheme()
```

The location of perfoming arts companies are more sparsely located around the city and less concentrated in Center city, as shown in Figure 3.4 below. 
```{r}
performingarts <- dat2 %>%
  filter(top_category == "Promoters of Performing Arts, Sports, and Similar Events" |
           top_category == "Performing Arts Companies") %>%
  st_transform('ESRI:102728') %>%
  mutate(Legend = "Performing Arts")

performingarts_net <-
  dplyr::select(performingarts) %>% 
  mutate(countPerformingarts = 1) %>% 
  aggregate(., fishnet, sum) %>%
  mutate(countPerformingarts = replace_na(countPerformingarts, 0),
         uniqueID = rownames(.),
         cvID = sample(round(nrow(fishnet) / 24), 
                       size=nrow(fishnet), replace = TRUE))
ggplot() +
  geom_sf(data = performingarts_net, aes(fill = countPerformingarts), color = NA) +
  scale_fill_viridis() +
  labs(title = "Count of Performing Arts Companies for Fishnet",
       subtitle = "Figure 3.4") +
  mapTheme()
```

Figure 3.5 below includes other sectors that contribute to Philadelphia's nightlife economy such as hotels and casinos. This figure combines the fishnets into a single panel that allows us to compare distribution of industries across business type.
```{r}
#Casinos
casinos <- dat2 %>%
  filter(top_category == "Gambling Industries") %>%
  st_transform('ESRI:102728') %>%
  mutate(Legend = "Casinos")

casinos_net <-
  dplyr::select(casinos) %>% 
  mutate(countCasinos = 1) %>% 
  aggregate(., fishnet, sum) %>%
  mutate(countCasinos = replace_na(countCasinos, 0),
         uniqueID = rownames(.),
         cvID = sample(round(nrow(fishnet) / 24), 
                       size=nrow(fishnet), replace = TRUE))

#Hotels
hotels <- dat2 %>%
  filter(top_category == "Traveler Accommodation") %>%
  st_transform('ESRI:102728') %>%
  mutate(Legend = "Hotels")

hotels_net <-
  dplyr::select(hotels) %>% 
  mutate(countHotels = 1) %>% 
  aggregate(., fishnet, sum) %>%
  mutate(countHotels = replace_na(countHotels, 0),
         uniqueID = rownames(.),
         cvID = sample(round(nrow(fishnet) / 24), 
                       size=nrow(fishnet), replace = TRUE))

# Combining fishnets into a single dataframe
vars_net <- 
  rbind(restaurants, bars, performingarts, casinos, hotels) %>%
  st_join(., fishnet, join=st_within) %>%
  st_drop_geometry() %>%
  group_by(uniqueID, Legend) %>%
  dplyr::summarize(count = n()) %>%
    full_join(fishnet) %>%
    spread(Legend,count, fill=0) %>%
    st_sf() %>%
    dplyr::select(-`<NA>`) %>%
    na.omit() %>%
    ungroup()

vars_net.long <- 
  gather(vars_net, Variable, value, -geometry, -uniqueID)

vars <- unique(vars_net.long$Variable)
mapList <- list()

#Plotting small multiple maps
for(i in vars){
  mapList[[i]] <- 
    ggplot() +
      geom_sf(data = filter(vars_net.long, Variable == i), aes(fill=value), colour=NA) +
      scale_fill_viridis(name="") +
      labs(title=i) +
      mapTheme()}

do.call(grid.arrange, c(mapList, ncol=3, top="Count of Nightlife Businesses per Fishnet", bottom = "Figure 3.5"))
```

## Nightlife hours

Next, we explore when visitors make trips to nightlife establishments. To do this, we worked with the popularity_by_hour SafeGraph variable, which sums the total number of visitors by hour for each month.

Figure 3.5 shows the average foot traffic by business type over the course of a day. This graphic indicates that though certain business types are considererd part of the nightlife economy, they do not exclusively experience traffic in the evenings. Restaurants are a good example of this, where the data indicates that the highest levels of traffic occur in the middle of the day. Arts venues, on the other hand, have a clear patterns indicating that they are more popular later in the day.
```{r traffic by nightlife establishment}
dat_hour %>%
  filter(top_category == "Drinking Places (Alcoholic Beverages)" |
           top_category == "Restaurants and Other Eating Places" |
           top_category == "Traveler Accommodation" |
           top_category == "Gambling Industries" |
           top_category == "Promoters of Performing Arts, Sports, and Similar Events" |
           top_category == "Performing Arts Companies") %>%
  group_by(Hour, top_category) %>%
  dplyr::summarize(Visitors = mean(Visitors)) %>%
  ggplot(., aes(x = Hour, y = Visitors)) + 
  geom_col() +
  labs(title = "Philadelphia Nightlife Organizations, Average Traffic by Hour",
       subtitle = "Figure 3.5") +
  facet_wrap(~top_category, scales = "free") +
  plotTheme()
```

The following animation shows restaurant visitor counts by census block group over the course of the day. We observe the highest amount of traffic in center city in the middle of the day and then a resurgence of traffic elsewhere in the city in the evening. This suggests that people dine at neighborhood establishments close to their home in the evening.
```{r}
dat_restaurants <-
  dat_hour %>%
  filter(top_category == "Restaurants and Other Eating Places") 

dat_restaurant_filter <-
  dat_restaurants %>%
  dplyr::rename(., GEOID10 = poi_cbg) %>%
  dplyr::group_by(GEOID10, Hour) %>%
  dplyr::summarize(Avg_Popularity = mean(Visitors),
            Total_Visits = sum(Visitors)) %>%
  left_join(phl_cbg) %>% 
  st_as_sf() %>%
  mutate(Visits_Per_Area = Total_Visits / Shape__Area * 100)

#Animation of restaurant popularity by hour
restaurant.animation.data <-
    dat_restaurant_filter %>%
    st_sf() %>%
    mutate(Pop_String = case_when(Visits_Per_Area < .4 ~ "5",
                              Visits_Per_Area >= .4 & Visits_Per_Area <.8 ~ "4",
                              Visits_Per_Area >= .8 & Visits_Per_Area <1.2 ~ "3",
                              Visits_Per_Area >= 1.2 & Visits_Per_Area <1.6 ~ "2",
                              Visits_Per_Area >= 2 ~ "1")) %>%
    mutate(Pop_String  = fct_relevel(Pop_String, "5","4","3","2","1"))

restaurant_animation <-
  ggplot() +
  geom_sf(data = phl_cbg, fill = "#440255", color = "transparent") +   
  geom_sf(data = restaurant.animation.data, aes(fill = Pop_String), color = "transparent") +
    scale_fill_manual(values = palette5) +
    labs(title = "Restaurant Popularity by Hour",
         subtitle = "One Hour Intervals: {current_frame}") +
  theme(panel.background = element_rect(fill = "black"),
         panel.grid.major = element_line(color = "transparent"),
          panel.grid.minor = element_line(colour = "transparent")) +
    transition_manual(Hour)
  

animate(restaurant_animation, duration=20, renderer = gifski_renderer())
```

The following animation shows the same metric for bars. We observe traffic increasing throughout the day starting in the afternoon.
```{r bar animation}
#Bar Analysis
#Filter Bars
dat_bars <- dat_hour %>%
  filter(top_category == "Drinking Places (Alcoholic Beverages)") 

#Merge CBGs with popularity by hour data
dat_bars_filter <-
  dat_bars %>%
  dplyr::rename(., GEOID10 = poi_cbg) %>%
  dplyr::group_by(GEOID10, Hour) %>%
  dplyr::summarize(Avg_Popularity = mean(Visitors),
            Total_Visits = sum(Visitors)) %>%
  left_join(phl_cbg) %>% 
  st_as_sf() %>%
  mutate(Visits_Per_Area = Total_Visits / Shape__Area * 100)

#Animation of bar popularity by hour
bar.animation.data <-
    dat_bars_filter %>%
    st_sf() %>%
    mutate(Pop_String = case_when(Visits_Per_Area == 0 ~ "5",
                              Visits_Per_Area > 0 & Visits_Per_Area <.2 ~ "4",
                              Visits_Per_Area >= .2 & Visits_Per_Area <.4 ~ "3",
                              Visits_Per_Area >= .4 & Visits_Per_Area <.6 ~ "2",
                              Visits_Per_Area >= .6 ~ "1")) %>%
    mutate(Pop_String  = fct_relevel(Pop_String, "5","4","3","2","1"))

bar_animation <-
  ggplot() +
  geom_sf(data = phl_cbg, fill = "#440255", color = "transparent", ) +
  geom_sf(data = bar.animation.data, aes(fill = Pop_String), color = "transparent") +
    scale_fill_manual(values = palette5) +
    labs(title = "Bar Popularity by Hour",
         subtitle = "One Hour Intervals: {current_frame}") +
  theme(panel.background = element_rect(fill = "black"),
         panel.grid.major = element_line(color = "transparent"),
          panel.grid.minor = element_line(colour = "transparent")) +
    transition_manual(Hour)

animate(bar_animation, duration=20, renderer = gifski_renderer())
```

## Nightlife visitors 

### Philadelphia visitor origins

Next we explore where visitors come from. Separated by nightlife establishment type, Figures 3.6a and 3.6b shows the average distance travelled from each census block group. The distance is calculated by measuring the euclidean distance between the centroid of each origin block group and the destination. Then we calculate the weighted mean of the distance travelled to account for destinations that people in a given block group visit more frequently.

For business types that are well-dispersed throughout the city, such as bars and restaurants, it appears that visitors typically travel shorter distances. For business types with fewer destinations, such as casinos, or that are clusters in specific areas of the city, such as hotels, it is common for visitors to make longer trips to these destinations.

Census block groups colored grey have no recorded visitors to the business type
```{r origin data wrangling}
#Preapring dataset to split out by individual cbgs
dat_cbg <- 
  dat2 %>%
  select(safegraph_place_id, 
         date_range_start, 
         top_category, 
         sub_category, 
         poi_cbg, 
         visitor_home_cbgs, 
         geometry) %>%
  mutate(visitor_home_cbgs = str_remove_all(visitor_home_cbgs, pattern = "\\[|\\]")) %>%
  mutate(visitor_home_cbgs = str_remove_all(visitor_home_cbgs, pattern = "\\{|\\}")) %>%
  mutate(visitor_home_cbgs = str_remove_all(visitor_home_cbgs, pattern = '\\"|\\"')) %>%
  mutate(visitor_home_cbgs = str_split(visitor_home_cbgs, pattern = ",")) %>%
  unnest(visitor_home_cbgs) %>%
  separate(.,
           visitor_home_cbgs,
           c("Visitor_CBG", "Visitors"),
           sep = ":") %>%
  mutate(Visitor_CBG = as.numeric(Visitor_CBG),
         poi_cbg = as.numeric(poi_cbg),
         Visitors = as.numeric(Visitors))

cbg_origin <- #takes a really long time!
  dat_cbg %>%
  filter(top_category == "Drinking Places (Alcoholic Beverages)" |
           top_category == "Restaurants and Other Eating Places" |
           top_category == "Traveler Accommodation" |
           top_category == "Gambling Industries" |
           top_category == "Promoters of Performing Arts, Sports, and Similar Events" |
           top_category == "Performing Arts Companies") %>% #filter for nightlife establishments
  st_drop_geometry() %>% #drop geometry to join with cbg file
  dplyr::rename(., GEOID10 = Visitor_CBG) %>% #renaming to match cbg file
  left_join(phl_cbg, by = "GEOID10") %>% # join to cbg file
  select(safegraph_place_id, 
         poi_cbg, 
         GEOID10, 
         Visitors, 
         geometry) %>% #clean up dataset
  rename(., cbg_origin = GEOID10, #clean up column names
         cbg_dest = poi_cbg,
         geometry_origin = geometry) %>%
  left_join(phila, by = "safegraph_place_id") %>% #join back to the SafeGraph locations
  select(safegraph_place_id, 
         top_category, 
         sub_category, 
         cbg_dest, 
         cbg_origin, 
         Visitors, 
         geometry_origin, 
         geometry) %>%
  rename(., geometry_dest = geometry) %>%
  drop_na(geometry_origin) %>% #removing geometries that didn't match (ie outside philily)
  mutate(distance = mapply(st_distance, st_centroid(geometry_origin), geometry_dest)) #calculate distance
```

```{r origin viz, message=FALSE, warning=FALSE, include=FALSE}
#Continuous
cbg_origin %>%
  group_by(cbg_origin, top_category) %>%
  summarize(avg_distance = mean(distance)) %>%  
  rename(., GEOID10 = cbg_origin) %>%
  left_join(phl_cbg, by = "GEOID10") %>%
  mutate(avg_distance = as.numeric(avg_distance)) %>%
  st_as_sf() %>%
  ggplot() +
  geom_sf(data = phl_cbg, fill = "grey70", color = "transparent") +
  geom_sf(aes(fill = avg_distance), color = "transparent") + 
  scale_fill_viridis() + 
  mapTheme() +
  labs(title = "How far do people travel to nightlife?", 
       subtitle = "Figure 3.6a") +
  facet_wrap(~top_category)

#Quintile
cbg_origin %>%
  drop_na(distance) %>%  
  group_by(cbg_origin, top_category) %>%
  summarize(avg_distance = weighted.mean(distance, Visitors)) %>%  
  rename(., GEOID10 = cbg_origin) %>%
  left_join(phl_cbg, by = "GEOID10") %>%
  mutate(avg_distance = as.numeric(avg_distance)) %>%
  st_as_sf() %>%
  ggplot() +
  geom_sf(data = phl_cbg, fill = "grey70", color = "transparent") +
  geom_sf(aes(fill = q5(avg_distance)), color = "transparent") + 
  scale_fill_manual(values = palette5,
                    aesthetics = c("colour", "fill"),
                    name = "Average Distance \n(Quintile)") +
  mapTheme() +
  labs(title = "How far do people travel to nightlife (quintile)?",
       subtitle = "Figure 3.6b") +
  facet_wrap(~top_category)
```

### Philadelphia destinations

Next, we study how far do Philadelphia visitors travel to nightlife establishemnts. These maps show the average distance travelled to each destination. Similar to the above, distance travelled is calculated by meauring the euclidena distance from each destination to the centroid of each visitor's block group. The metric shown here represented the weighted average of trips and distances.

In general, we see that visitors travel further to Center City destinations. Importantly, these graphics exclude all trips that originate outside of Philadelphia.

```{r destination data wrangling}
cbg_dest <-
  dat_cbg %>%
  filter(top_category == "Drinking Places (Alcoholic Beverages)" |
           top_category == "Restaurants and Other Eating Places" |
           top_category == "Traveler Accommodation" |
           top_category == "Gambling Industries" |
           top_category == "Promoters of Performing Arts, Sports, and Similar Events" |
           top_category == "Performing Arts Companies") %>%
  st_drop_geometry() %>%
  rename(., GEOID10 = Visitor_CBG) %>%
  left_join(phl_cbg, by = "GEOID10") %>%
  select(safegraph_place_id, 
         GEOID10, 
         poi_cbg, 
         Visitors, 
         geometry) %>%
  rename(., cbg_origin = GEOID10,
         geometry_origin = geometry,
         cbg_dest = poi_cbg) %>%
  left_join(phila, by = "safegraph_place_id") %>%
  select(safegraph_place_id, 
         top_category, 
         sub_category, 
         cbg_origin, 
         cbg_dest, 
         Visitors, 
         geometry_origin, 
         geometry) %>%
  rename(., geometry_dest = geometry) %>%
  drop_na(geometry_origin) %>% #dropping origins outside of philadelphia
  mutate(distance = mapply(st_distance, st_centroid(geometry_origin), st_centroid(geometry_dest)))
```

```{r destination viz}
#Continuous
cbg_dest %>%
  group_by(safegraph_place_id, top_category) %>%
  summarize(avg_distance = weighted.mean(distance, Visitors)) %>%
  dplyr::select(safegraph_place_id, avg_distance) %>%
  left_join(phila, by = "safegraph_place_id") %>%
  st_as_sf() %>%
  ggplot() +
  geom_sf(data = phl_cbg, fill = "grey70", color = "transparent") +
  geom_sf(aes(color = avg_distance), size = 1) + 
  scale_fill_viridis(aesthetics = "color") +
  mapTheme() +
  labs(title = "Which nightlife destinations do visitors travel the furthest?",
       subtitle = "Figure 3.7a") +
  facet_wrap(~top_category)

#Quintile
cbg_dest %>%
  group_by(safegraph_place_id, top_category) %>%
  summarize(avg_distance = weighted.mean(distance, Visitors)) %>%
  dplyr::select(safegraph_place_id, avg_distance) %>%
  left_join(phila, by = "safegraph_place_id") %>%
  st_as_sf() %>%
  ggplot() +
  geom_sf(data = phl_cbg, fill = "grey70", color = "transparent") +
  geom_sf(aes(color = q5(avg_distance)), size = 1) + 
  scale_fill_manual(values = palette5,
                    aesthetics = c("colour", "fill"),
                    name = "Average Distance \n(Quintile)") +
  mapTheme() +
  labs(title = "Which nightlife destinations do visitors travel the farthest?",
       subtitle = "Figure 3.7b") +
  facet_wrap(~top_category)
```

## Nightlife corridors in Philadelphia

Next, we observe foot traffic along Philadelphia's commercial corridors. This analysis relies on a shapefile from the City of Philadelphia's Planning department which demarcates individual corridors and districts throughout the city. According to the [available metadata](https://metadata.phila.gov/#home/datasetdetails/564236a55737e1f263ae5e3f/representationdetails/56423a4e902dbdd813db9a55/) "locations range from large, regional and specialty destinations to corridors that reflect the evolving economy, culture, and aesthetic traditions of surrounding neighborhoods." This means that the gegoraphies vary in size and character.

Figure 3.8 below maps the commercial corridors and shows the corridor-wide vacancy rate, which varies from close to 0% to upwards of 60% throughout the city. Smaller corridors in far West Philadelphia, North Philadelphia tend to have the highest vacancy rates.
```{r}
###Corridors
#Load Philadelphia commercial corridors dataset and transform
corridors <- phl_corridors

dat_restaurants <-
dat_hour %>%
  filter(top_category == "Restaurants and Other Eating Places") 

corridors_filter <- corridors %>%
  select(OBJECTID, NAME, GLA, P_DIST, ST_EXT, PT_ADD, VAC_RATE) %>%
  mutate(VAC_RATE = str_remove_all(VAC_RATE, pattern = "%")) %>%
  mutate(VAC_RATE = as.numeric(VAC_RATE)) %>%
  st_as_sf() 
# %>%
#   st_transform('ESRI:102271') 
  
# Plot commercial corridors with colors as vacancy rate
ggplot() +
  geom_sf(data = phillyBoundary, fill = "black") +
  geom_sf(data = corridors_filter, aes(fill = VAC_RATE), color = "transparent") +
  scale_fill_viridis_c(option = "plasma") +
  labs(title = "Philadelphia Commercial Corridors",
       subtitle = "Figure 3.8") + 
  mapTheme()

```

Figure 3.9 below breaks down the count of restaunts in a given corridor per square miile.  We find that central corridors and districts tend to have more restaurants per square mile. Other corridors. That said, corridors across the city have a high concentration of restaurants as well.
```{r}
###Restaurants (takes a really long time)
dat_restaurants_grid <- dat_restaurants %>%
  select(geometry) %>%
  na.omit() %>%
  st_as_sf() %>%
  #st_transform('ESRI:102271') %>% 
    distinct()

#Merge restaurants with commercial corridors 
corridors_restaurants <- 
  select(dat_restaurants_grid) %>% 
  mutate(countRestaurant = 1) %>% 
  aggregate(., corridors_filter, sum) %>%
  mutate(countRestaurant = replace_na(countRestaurant, 0),
         uniqueID = rownames(.),
         area = st_area(geometry) * .00000038610,
         count_per_mile = as.numeric(countRestaurant / area),
         cvID = sample(round(nrow(corridors_filter) / 24), size=nrow(corridors_filter), replace = TRUE))

#Plot map of number of restaurants per square mile
ggplot() +
  geom_sf(data = phillyBoundary, fill = "black") +
  geom_sf(data = corridors_restaurants, aes(fill = count_per_mile), color = "transparent") +
  scale_fill_viridis(trans = "sqrt") +
  labs(title = "Count of Restaurants per Square Mile in Each Commercial Corridor",
       subtitle = "Figure 3.9") +
  mapTheme()
```

Figure 3.10 shows the number of bars in each commercial corridor per square mile. Again, we see the corridors around Center City generally showing a higher concentration of bars.
```{r Bars by corridor}
###Bars
dat_bars_grid <- dat_bars %>%
  select(geometry) %>%
  #na.omit() %>%
  st_as_sf() %>%
  # st_transform('ESRI:102271') %>% 
    distinct()
  
corridors_bars <- 
  dplyr::select(dat_bars_grid) %>% 
  mutate(countBar = 1) %>% 
  aggregate(., corridors_filter, sum) %>%
  mutate(countBar = replace_na(countBar, 0),
         uniqueID = rownames(.),
         area = st_area(geometry) * .00000038610,
         count_per_mile = as.numeric(countBar / area),
         cvID = sample(round(nrow(corridors_filter) / 24), size=nrow(corridors_filter), replace = TRUE))

#Plot map of number of bars per square mile
ggplot() +
  geom_sf(data = phillyBoundary, fill = "black") +
  geom_sf(data = corridors_bars, aes(fill = count_per_mile), color = "transparent") +
  scale_fill_viridis(trans = "sqrt") +
  labs(title = "Count of Bars per Square Mile in Each Commercial Corridor",
       subtitle = "Figure 3.10") +
  mapTheme()
```

## Neighborhood and corridor comparisons

Finally, we began exploring businesses at the corridor level and understand how different corridors attract different visitors from across the city. We plan to systematically expand the scope of this study to other corridors across the city. We also believe that this data is particularly relevant to business and economic development professionals in the COVID-19 recovery efforts.

Using the Commercial Corridor shapefile available on Open Data Philly, we selected two corridors to compare:  East Girard Avenue, a central corridor in the Fishtown neighborhood, and West Girard Avenue in Brewerytown. 

```{r East Girard data wrangling}
#East Girard Corridor
EastGirard_corr <- 
  phl_corridors %>% 
  filter(NAME == "East Girard") %>% 
  st_as_sf()

dat_EGC <- dat2[EastGirard_corr,]

dat_EGC_cbg <- 
  dat_EGC %>%
  select(safegraph_place_id, 
         date_range_start, 
         top_category, 
         sub_category, 
         visitor_home_cbgs, 
         geometry) %>%
  mutate(visitor_home_cbgs = str_remove_all(visitor_home_cbgs, pattern = "\\[|\\]")) %>%
  mutate(visitor_home_cbgs = str_remove_all(visitor_home_cbgs, pattern = "\\{|\\}")) %>%
  mutate(visitor_home_cbgs = str_remove_all(visitor_home_cbgs, pattern = '\\"|\\"')) %>%
  mutate(visitor_home_cbgs = str_split(visitor_home_cbgs, pattern = ",")) %>%
  unnest(visitor_home_cbgs) %>%
  separate(.,
           visitor_home_cbgs,
           c("cbg_origin", "visitors"),
           sep = ":") %>%
  mutate(cbg_origin = as.numeric(cbg_origin),
         visitors = as.numeric(visitors)) %>%
  filter(top_category == "Drinking Places (Alcoholic Beverages)" |
           top_category == "Restaurants and Other Eating Places" |
           top_category == "Traveler Accommodation" |
           top_category == "Gambling Industries" |
           top_category == "Promoters of Performing Arts, Sports, and Similar Events" |
           top_category == "Performing Arts Companies") %>%
    mutate(corridor = "East Girard")

#West Girard
WestGirard_corr <- 
  phl_corridors %>% 
  filter(NAME == "West Girard") %>% 
  st_as_sf()

dat_WGC <- dat2[WestGirard_corr,]

dat_WGC_cbg <- 
  dat_WGC %>%
  select(safegraph_place_id, 
         date_range_start, 
         top_category, 
         sub_category, 
         visitor_home_cbgs, 
         geometry) %>%
  mutate(visitor_home_cbgs = str_remove_all(visitor_home_cbgs, pattern = "\\[|\\]")) %>%
  mutate(visitor_home_cbgs = str_remove_all(visitor_home_cbgs, pattern = "\\{|\\}")) %>%
  mutate(visitor_home_cbgs = str_remove_all(visitor_home_cbgs, pattern = '\\"|\\"')) %>%
  mutate(visitor_home_cbgs = str_split(visitor_home_cbgs, pattern = ",")) %>%
  unnest(visitor_home_cbgs) %>%
  separate(.,
           visitor_home_cbgs,
           c("cbg_origin", "visitors"),
           sep = ":") %>%
  mutate(cbg_origin = as.numeric(cbg_origin),
         visitors = as.numeric(visitors)) %>%
  filter(top_category == "Drinking Places (Alcoholic Beverages)" |
           top_category == "Restaurants and Other Eating Places" |
           top_category == "Traveler Accommodation" |
           top_category == "Gambling Industries" |
           top_category == "Promoters of Performing Arts, Sports, and Similar Events" |
           top_category == "Performing Arts Companies") %>%
  mutate(corridor = "West Girard")
```

Figures 3.11a and 3.11b show the volume of visitors to the East Girard Corridor (outlined in red) by census block group. While visitors from across the city frequent East Girard corridor, particularly for its restaurants, Figure 3.11b clarifies that the corridor mostly caters to a local customer base of surrounding census tracts.

Block groups shaded gray indicate that no resident made a trip to this corridor.
```{r}
dat_EGC_cbg %>%
  dplyr::group_by(cbg_origin, top_category) %>%
  dplyr::summarize(visitors = sum(visitors)) %>%
  st_drop_geometry() %>%
  dplyr::rename(., GEOID10 = cbg_origin) %>%
  left_join(phl_cbg, by = "GEOID10") %>%
  st_as_sf() %>%
  drop_na(geometry) %>% 
  ggplot() +
  geom_sf(data = phl_cbg, fill = "grey70", color = "transparent") +
  geom_sf(aes(fill = q5(visitors), geometry = geometry), color = "transparent") + 
  geom_sf(data = phl_nhoods, color = "white", fill = "transparent") +
  geom_sf(data = EastGirard_corr, color = "red", fill = "transparent", lwd = 1) +
  scale_fill_manual(values = palette5,
                    aesthetics = c("colour", "fill"),
                    name = "Visitor Count \n(Quintile)") +
  mapTheme() +
  labs(title = "East Girard Restaurants: Where do visitors come from?",
       subtitle = "Figure 3.11a") +
  facet_wrap(~top_category)

dat_EGC_cbg %>%
  dplyr::group_by(cbg_origin, top_category) %>%
  dplyr::summarize(visitors = sum(visitors)) %>%
  st_drop_geometry() %>%
  dplyr::rename(., GEOID10 = cbg_origin) %>%
  left_join(phl_cbg, by = "GEOID10") %>%
  st_as_sf() %>%
  drop_na(geometry) %>% 
  ggplot() +
  geom_sf(data = phl_cbg, fill = "grey70", color = "transparent") +
  geom_sf(aes(fill = visitors, geometry = geometry), color = "transparent") + 
  geom_sf(data = phl_nhoods, color = "white", fill = "transparent") +
  geom_sf(data = EastGirard_corr, color = "red", fill = "transparent", lwd = .5) +
  scale_fill_viridis() +
  mapTheme() +
  labs(title = "East Girard Restaurants: Where do visitors come from?",
       subtitle = "Figure 3.8b") +
  facet_wrap(~top_category)
```

Figures 3.12a and 3.12b show the volume of visitors to West Girard corridor. While the corridor appears to draw visitors from fewer census tracts overall, the corridor draws from a similarly local customer base of the surrounding block groups.

As before, block groups shaded gray have no residents that made trips to this corridor.
```{r West Girard viz}
dat_WGC_cbg %>%
  dplyr::group_by(cbg_origin, top_category) %>%
  dplyr::summarize(visitors = sum(visitors)) %>%
  st_drop_geometry() %>%
  dplyr::rename(., GEOID10 = cbg_origin) %>%
  left_join(phl_cbg, by = "GEOID10") %>%
  st_as_sf() %>%
  drop_na(geometry) %>% 
  ggplot() +
  geom_sf(data = phl_cbg, fill = "grey70", color = "transparent") +
  geom_sf(aes(fill = q5(visitors), geometry = geometry), color = "transparent") + 
  geom_sf(data = phl_nhoods, color = "white", fill = "transparent") +
  geom_sf(data = WestGirard_corr, color = "red", fill = "transparent", lwd = 1) +
  scale_fill_manual(values = palette5,
                    aesthetics = c("colour", "fill"),
                    name = "Visitor Count \n(Quintile)") +
  mapTheme() +
  labs(title = "West Girard: Where do visitors come from?",
       subtitle = "Figure 3.12a") +
  facet_wrap(~top_category)

dat_WGC_cbg %>%
  dplyr::group_by(cbg_origin, top_category) %>%
  dplyr::summarize(visitors = sum(visitors)) %>%
  st_drop_geometry() %>%
  dplyr::rename(., GEOID10 = cbg_origin) %>%
  left_join(phl_cbg, by = "GEOID10") %>%
  st_as_sf() %>%
  drop_na(geometry) %>% 
  ggplot() +
  geom_sf(data = phl_cbg, fill = "grey70", color = "transparent") +
  geom_sf(aes(fill = visitors, geometry = geometry), color = "transparent") + 
  geom_sf(data = phl_nhoods, color = "white", fill = "transparent") +
  geom_sf(data = WestGirard_corr, color = "red", fill = "transparent", lwd = .5) +
  scale_fill_viridis() +
  mapTheme() +
  labs(title = "East Girard Restaurants: Where do visitors come from?",
       subtitle = "Figure 3.12b") +
  facet_wrap(~top_category)
```

We can further explore the individual corridor analysis by tying the trip origins to census data. This allows us to construct a customer profile of the people visiting specific corridors throughout the city.
```{r census pull}
phl_blockgroups <- 
  get_acs(geography = "block group", 
        variables = c("B01003_001E", 
                      "B02001_002E", 
                      "B01002_001E",
                      "B19013_001E", 
                      "B25064_001E"),
        year=2018, 
        state=42, 
        county=101, 
        geometry=T, 
        output = "wide") %>%
  st_transform('ESRI:102728')%>%
  rename(TotalPop = B01003_001E,
         Whites = B02001_002E,
         MedAge = B01002_001E,
         MedHHInc = B19013_001E,
         MedRent = B25064_001E,
         GEOID10 = GEOID) %>%
  dplyr::select(-NAME, -starts_with("B")) %>%
  mutate(pctWhite = ifelse(TotalPop > 0, Whites / TotalPop,0),
         GEOID10 = as.numeric(GEOID10)) %>%
  dplyr::select(-Whites) 
```

The below Figure 3.13 compares demographic indicators pulled from the census across the two customer bases. East Girard corridor's customer base, shown in purple, tend to come from whiter census tracts with a higher median household income and median rent. West Girard corridor's customer base tend to come from less white census tracts with lower incomes and median rent. Both corridors, however, pull from census tracts with a similar median age.
```{r corridor comparison}
rbind(dat_EGC_cbg, dat_WGC_cbg) %>%
  rename(GEOID10 = cbg_origin) %>%
  dplyr::select(GEOID10, visitors, corridor) %>%
  st_drop_geometry() %>%
  left_join(phl_blockgroups, by = "GEOID10") %>%
  pivot_longer(.,
               cols = c("MedAge", "MedHHInc", "MedRent", "pctWhite"),
               names_to = "variable") %>%
  dplyr::select(GEOID10, corridor, visitors, variable, value) %>%
  ggplot(., aes(x = value, fill = corridor, weight = visitors)) + 
  geom_density(alpha = .5) +
  scale_fill_manual(values = palette3,
                    aesthetics = c("colour", "fill"),
                    name = "Visitor Count \n(Quintile)") +
  facet_wrap( ~ variable, scales = "free") +
  plotTheme() +
  labs(title = 'East Girard vs. West Girard Corridors',
       subtitle = "Figure 3.13")
```

# 4. Next Steps

We will continue our Exploratory Data Analysis and plan to focus on the following topics:

* Defining and demarcating distinct Philadelphia nightlife corridors.
* Expanding analysis to the regional scale.  Currently we only review trips originating within Phialdelphia, but it will be important to understand which districts have a more regional approach.
* Understand how COVID-19 has impacted nightlife foot traffic in Philadelphia? Have corridors been impacted evenly?